---
title: "R Notebook"
output: html_notebook
---

#### Project:
HTP

#### Analyst (Email):
Jessica Shaw (jessica.shaw@cuanschutz.edu)

#### Goal:
Analysis of MSD cytokine concentrations in association with AD biomarkers, as measured by 4-plex SIMOA.

Baseline linear models were be fitted for the log2 concentration of each of four AD biomarkers (NfL, GFAP, TAU, and UCH-L1), with independent variables T21, Age, and Female sex.

Then, the linear models were re-fitted with additional adjustment for each of 54 cytokines.

This analysis provides estimates of association between cytokine concentrations and levels of AD biomarkers. Furthermore, we can infer whether cytokine levels mediate the association between T21 and AD biomarker levels (this part isn't shown yet in this report).

```{r}
#rm(list = ls(all = TRUE))
#library(openxlsx)
library(data.table)
library(dplyr)
library(tidyr)
library(broom)
#library(lattice)
library(ggplot2)
#library(infer)
#library(miceadds)
#library(FedData)
#library(xlsx)
library(ggrepel)
library(reshape2)
library(tictoc)
library(car)
library(broomExtra)
library(tibble)
```

```{r}
#setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/Pairwise_Omics/MSD_SIMOA/Data")
#simoa_msd.complete<-fread("PairwiseOmics_MSD_SIMOA_073020_SIMOAwide_MSDlong_MatchLabID_v2.1_JRS.csv")

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/Pairwise_Omics/MSD_SIMOA/Data")
simoa_msd.complete<-fread("PairwiseOmics_MSD_SIMOA_080620_SIMOAwide_MSDlong_MatchLabID_v2.2_JRS.csv")

simoa_msd.complete
```

#### Get summary statistics for each AD biomarker
```{r}
summaryStats.SIMOA<-simoa_msd.complete %>%
  select(RecordID, LabID, NfL, GFAP, TAU, `UCH-L1`) %>%
  gather(key="SIMOA_Biomarker", value="SIMOA_Concentration", NfL:`UCH-L1`) %>%
  group_by(SIMOA_Biomarker) %>%
  summarise(N=n(),
            Min = min(SIMOA_Concentration),
            p01 = quantile(SIMOA_Concentration, 0.01),
            p05 = quantile(SIMOA_Concentration, 0.05),
            Mean = mean(SIMOA_Concentration),
            Med = median(SIMOA_Concentration),
            Max = max(SIMOA_Concentration))
summaryStats.SIMOA
# UCH-L1 has a minimum value of 0, so will require an offset for log2 transformation.
```

#### MEDIATION ANALYSIS
Goal: Identify which cytokines mediate the association of T21 and AD biomarkers levels.

##### Mediation Analysis | Step 1: Run models for only AD biomarkers in association with age, T21, * in THIS subset of data *
```{r}
set.seed(1234)

# LM without random effect for Source:
baselineLM.NfL<-lm(log2(NfL) ~ T21 + Age + Female, data=simoa_msd.complete)
baselineLM.GFAP<-lm(log2(GFAP) ~ T21 + Age + Female, data=simoa_msd.complete)
baselineLM.TAU<-lm(log2(TAU) ~ T21 + Age + Female, data=simoa_msd.complete)
baselineLM.UCHL1<-lm(log2(`UCH-L1` + 0.00001) ~ T21 + Age + Female, data=simoa_msd.complete)

# LMM with random effect for Source:
baselineLMM.NfL<-lmerTest::lmer(log2(NfL) ~ T21 + Age + Female + (1|Source), REML=FALSE, data=simoa_msd.complete)
baselineLMM.GFAP<-lmerTest::lmer(log2(GFAP) ~ T21 + Age + Female + (1|Source), REML=FALSE, data=simoa_msd.complete)
baselineLMM.TAU<-lmerTest::lmer(log2(TAU) ~ T21 + Age + Female + (1|Source), REML=FALSE, data=simoa_msd.complete)
baselineLMM.UCHL1<-lmerTest::lmer(log2(`UCH-L1` + 0.00001) ~ T21 + Age + Female + (1|Source), REML=FALSE, data=simoa_msd.complete)

AIC(baselineLM.NfL, baselineLMM.NfL,
    baselineLM.GFAP, baselineLMM.GFAP,
    baselineLM.TAU, baselineLMM.TAU,
    baselineLM.UCHL1, baselineLMM.UCHL1)
# NfL improves dramatically with random intercept for Source. GFAP improves somewhat. TAU and UCH-L1 don't change much - very slightly worse fit with random effect...

simoa_msd.complete %>%
  select(RecordID, Source) %>%
  unique() %>%
  group_by(Source) %>%
  summarise(N=n())

simoa_msd.complete %>%
  select(RecordID, Source, T21) %>%
  unique() %>%
  group_by(Source, T21) %>%
  summarise(N=n())

# Only N=3 from NDSC 2019, so in this particular case we really may not be able to trust a random effect... hmmm...
```

```{r}
tidyBasic_BaselineLMM.NfL<-baselineLMM.NfL %>%
  tidy() %>%
  mutate(AD_Biomarker = "NfL",
         Cytokine = "NONE",
         Model = "log2(NfL) ~ T21 + Age + Female") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLMM.GFAP<-baselineLMM.GFAP %>%
  tidy() %>%
  mutate(AD_Biomarker = "GFAP",
         Cytokine = "NONE",
         Model = "log2(GFAP) ~ T21 + Age + Female") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLMM.TAU<-baselineLMM.TAU %>%
  tidy() %>%
  mutate(AD_Biomarker = "TAU",
         Cytokine = "NONE",
         Model = "log2(TAU) ~ T21 + Age + Female") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLMM.UCHL1<-baselineLMM.UCHL1 %>%
  tidy() %>%
  mutate(AD_Biomarker = "UCH-L1",
         Cytokine = "NONE",
         Model = "log2(`UCH-L1` + offset.UCHL1) ~ T21 + Age + Female") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

# tidyBasic_BaselineLMM.NfL
# tidyBasic_BaselineLMM.GFAP
# tidyBasic_BaselineLMM.TAU
# tidyBasic_BaselineLMM.UCHL1

ageResultsBasic_BASELINElmm.NfL<-tidyBasic_BaselineLMM.NfL %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElmm.GFAP<-tidyBasic_BaselineLMM.GFAP %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElmm.TAU<-tidyBasic_BaselineLMM.TAU %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElmm.UCHL1<-tidyBasic_BaselineLMM.UCHL1 %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))

T21resultsBasic_BASELINElmm.NfL<-tidyBasic_BaselineLMM.NfL %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElmm.GFAP<-tidyBasic_BaselineLMM.GFAP %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElmm.TAU<-tidyBasic_BaselineLMM.TAU %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElmm.UCHL1<-tidyBasic_BaselineLMM.UCHL1 %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))


# LM without random effect for Source:
tidyBasic_BaselineLM.NfL<-baselineLM.NfL %>%
  tidy() %>%
  mutate(AD_Biomarker = "NfL",
         Cytokine = "NONE",
         Model = "log2(NfL) ~ T21 + Age + Female + (1|Source)") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLM.GFAP<-baselineLM.GFAP %>%
  tidy() %>%
  mutate(AD_Biomarker = "GFAP",
         Cytokine = "NONE",
         Model = "log2(GFAP) ~ T21 + Age + Female + (1|Source)") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLM.TAU<-baselineLM.TAU %>%
  tidy() %>%
  mutate(AD_Biomarker = "TAU",
         Cytokine = "NONE",
         Model = "log2(TAU) ~ T21 + Age + Female + (1|Source)") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

tidyBasic_BaselineLM.UCHL1<-baselineLM.UCHL1 %>%
  tidy() %>%
  mutate(AD_Biomarker = "UCH-L1",
         Cytokine = "NONE",
         Model = "log2(`UCH-L1` + offset.UCHL1) ~ T21 + Age + Female + (1|Source)") %>%
  select(AD_Biomarker, Cytokine, Model, everything()) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "df", "P"))

# tidyBasic_BaselineLM.NfL
# tidyBasic_BaselineLM.GFAP
# tidyBasic_BaselineLM.TAU
# tidyBasic_BaselineLM.UCHL1

ageResultsBasic_BASELINElm.NfL<-tidyBasic_BaselineLM.NfL %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElm.GFAP<-tidyBasic_BaselineLM.GFAP %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElm.TAU<-tidyBasic_BaselineLM.TAU %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))
ageResultsBasic_BASELINElm.UCHL1<-tidyBasic_BaselineLM.UCHL1 %>%
  filter(Term=="Age") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_Age", "SE_Age", "Statistic_Age", "df_Age", "P_Age"))

T21resultsBasic_BASELINElm.NfL<-tidyBasic_BaselineLM.NfL %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElm.GFAP<-tidyBasic_BaselineLM.GFAP %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElm.TAU<-tidyBasic_BaselineLM.TAU %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))
T21resultsBasic_BASELINElm.UCHL1<-tidyBasic_BaselineLM.UCHL1 %>%
  filter(Term=="T21") %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate_T21", "SE_T21", "Statistic_T21", "df_T21", "P_T21"))

# View the results:

T21resultsBasic_BASELINElmm.NfL
T21resultsBasic_BASELINElm.NfL

T21resultsBasic_BASELINElmm.GFAP
T21resultsBasic_BASELINElm.GFAP

T21resultsBasic_BASELINElmm.TAU
T21resultsBasic_BASELINElm.TAU

T21resultsBasic_BASELINElmm.UCHL1
T21resultsBasic_BASELINElm.UCHL1

# ageResultsBasic_BASELINElmm.NfL
# ageResultsBasic_BASELINElmm.GFAP
# ageResultsBasic_BASELINElmm.TAU
# ageResultsBasic_BASELINElmm.UCHL1
# ageResultsBasic_BASELINElm.NfL
# ageResultsBasic_BASELINElm.GFAP
# ageResultsBasic_BASELINElm.TAU
# ageResultsBasic_BASELINElm.UCHL1

# All four are highly significantly associated with T21, regardless of whether we include random intercept for Source. In general, including the random intercept for Source makes T21 much more significant.

# Decision: Only for proof of concept, use only fixed-effects models for this script. Come back later to figure out how to deal with random effects in the context of mediation analysis.
```

##### Mediation Analysis | Step 2: Run models for only cytokines in association with age, T21, * in THIS subset of data *
```{r}
set.seed(1234)

msd_only<-simoa_msd.complete %>% select(-c(NfL, GFAP, TAU, `UCH-L1`)) %>% unique() %>%
  mutate(Index_Analyte = as.numeric(as.factor(MSD_Analyte)))
msd_only

lm.Cytokine<-list()
tidy.Cytokine<-list()
for ( i in 1:max(msd_only$Index_Analyte) ) {

  df<-msd_only %>% filter(Index_Analyte==i)
  msd_analyte<-df$MSD_Analyte[1]
  
  lm.Cytokine[[i]]<-lm(log2(MSD_Concentration) ~ T21 + Age + Female, data = df)
  tidy.Cytokine[[i]]<-lm.Cytokine[[i]] %>% tidy() %>% filter(term!="(Intercept)") %>%
    mutate(`AD Biomarker` = "NONE", Cytokine = msd_analyte,
           Model = "log2(MSD_Concentration) ~ T21 + Age + Female") %>%
    select(`AD Biomarker`, Cytokine, Model, everything())
  
}

T21results.Cytokine<-tidy.Cytokine %>%
  rbindlist() %>%
  filter(term=="T21") %>%
  select(-term) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Estimate_T21", "SE_T21", "Statistic_T21",
                 "P_T21")) %>%
  mutate(Sig_Cytokine = ifelse(P_T21<0.05, 1, 0))

T21results.Cytokine %>%
  arrange(P_T21)
```

### Run cytokine-adjusted models

Above we fitted a preliminary model with adjustment for standard covariates (T21, age, and sex) and batch effect correction for sample source. We'll refer to this model specification as the 'basic' model, specified by the formula:
  log2(Conc. AD Biomarker) ~ T21 + Age + Sex + (1|Source)

We also identified the best-fitting model specification for our AD biomarkers. We'll refer to this model specification as the 'optimal' model, specified by the formula:
  log2(Conc. AD Biomarker) ~ T21 + Age + Sex + (T21 x Age) + (T21 x Sex) + (1|Source)

The code below re-fits the 'basic' and 'optimal' models with additional adjustment for the level of one cytokine at a time (cytokine_i). In other words, we fit one mixed model for each unique pair of AD biomarker (measured by SIMOA) and cytokine (measured by MSD):
  * Basic model + Cytokine_i
  * Optimal model + Cytokine_i

##### Mediation Analysis | Step 3: Run cytokine-adjusted models
```{r}
# Prepare an analysis dataset with an index for cytokine, to facilitate adding one cytokine at a time to each model:
simoa_msd.complete01<-simoa_msd.complete %>%
  select(MSD_Analyte) %>%
  unique() %>%
  mutate(Index_MSD = seq(1:length(MSD_Analyte))) %>%
  full_join(simoa_msd.complete, by="MSD_Analyte")
#simoa_msd.complete01

# Define objects to store the various outputs produced in the for loop below:
df<-list()
cytokine_i<-list() # Where i takes on values 1 through 49 for the 1st through 49th cytokine in the joint SIMOA x MSD dataset.
lmAdjCytokine.NfL<-list()
lmAdjCytokine.GFAP<-list()
lmAdjCytokine.TAU<-list()
lmAdjCytokine.UCHL1<-list()
tidyAdjCytokine.NfL<-list()
tidyAdjCytokine.GFAP<-list()
tidyAdjCytokine.TAU<-list()
tidyAdjCytokine.UCHL1<-list()
vifAdjCytokine.NfL<-list()
vifAdjCytokine.GFAP<-list()
vifAdjCytokine.TAU<-list()
vifAdjCytokine.UCHL1<-list()

set.seed(1234)

for (i in 1:max(simoa_msd.complete01$Index_MSD)) {
  
  df[[i]]<-simoa_msd.complete01 %>%
    filter(Index_MSD == i) %>%
    select(RecordID, NfL, GFAP, TAU, `UCH-L1`, MSD_Analyte, MSD_Concentration, T21, Age, Female, Source) %>%
    unique() %>%
    na.omit()
  
  cytokine_i[[i]]<-df[[i]]$MSD_Analyte[1]
  colnames(df[[i]])[7]<-cytokine_i[[i]]

  # Template for 'AdjCytokine' model + Cytokine = The model for log2(Conc) of AD biomarker, with adjustment only for T21, age, sex, and source.
  formula_adjCytokine<-paste("log2(`", "placeholder.AD_Biomarker", "` + offset) ~ log2(`",
                                   colnames(df[[i]])[7], # column for cytokine concentration
                                   "`) + T21 + Age + Female",
                                   sep = "")

  ###########
  ##  NfL  ##
  ###########
  ADbiomarker<-"NfL"
  offset<-0
  formula_adjCytokine_i<-as.formula( gsub("placeholder.AD_Biomarker", ADbiomarker, formula_adjCytokine) )

  # 'AdjCytokine' model + Cytokine:
  lmAdjCytokine.NfL[[i]]<-try( lm( formula_adjCytokine_i, data=df[[i]] ) )
  tidyAdjCytokine.NfL[[i]]<-try( lmAdjCytokine.NfL[[i]] %>%
                             tidy() %>%
                             mutate(AD_Biomarker = ADbiomarker,
                                    Cytokine = cytokine_i[[i]],
                                    Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                             select(AD_Biomarker, Cytokine, Model, everything()) )
  vifAdjCytokine.NfL[[i]]<-try( lmAdjCytokine.NfL[[i]] %>%
                       car::vif() %>%
                       as.data.frame() %>%
                       rownames_to_column("Model Term") %>%
                       `colnames<-`(c("Model Term", "VIF(Model Term)")) %>%
                       mutate(Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                       select(Model, `Model Term`, `VIF(Model Term)`) )
  

   ############
  ##  GFAP  ##
  ############
  ADbiomarker<-"GFAP"
  offset<-0
  formula_adjCytokine_i<-as.formula( gsub("placeholder.AD_Biomarker", ADbiomarker, formula_adjCytokine) )

  # 'AdjCytokine' model + Cytokine:
  lmAdjCytokine.GFAP[[i]]<-try( lm( formula_adjCytokine_i, data=df[[i]] ) )
  tidyAdjCytokine.GFAP[[i]]<-try( lmAdjCytokine.GFAP[[i]] %>%
                             tidy() %>%
                             mutate(AD_Biomarker = ADbiomarker,
                                    Cytokine = cytokine_i[[i]],
                                    Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                             select(AD_Biomarker, Cytokine, Model, everything()) )
  vifAdjCytokine.GFAP[[i]]<-try( lmAdjCytokine.GFAP[[i]] %>%
                       car::vif() %>%
                       as.data.frame() %>%
                       rownames_to_column("Model Term") %>%
                       `colnames<-`(c("Model Term", "VIF(Model Term)")) %>%
                       mutate(Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                       select(Model, `Model Term`, `VIF(Model Term)`) )
  

   ###########
  ##  TAU  ##
  ###########
  ADbiomarker<-"TAU"
  offset<-0
  formula_adjCytokine_i<-as.formula( gsub("placeholder.AD_Biomarker", ADbiomarker, formula_adjCytokine) )

  # 'AdjCytokine' model + Cytokine:
  lmAdjCytokine.TAU[[i]]<-try( lm( formula_adjCytokine_i, data=df[[i]] ) )
  tidyAdjCytokine.TAU[[i]]<-try( lmAdjCytokine.TAU[[i]] %>%
                             tidy() %>%
                             mutate(AD_Biomarker = ADbiomarker,
                                    Cytokine = cytokine_i[[i]],
                                    Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                             select(AD_Biomarker, Cytokine, Model, everything()) )
  vifAdjCytokine.TAU[[i]]<-try( lmAdjCytokine.TAU[[i]] %>%
                       car::vif() %>%
                       as.data.frame() %>%
                       rownames_to_column("Model Term") %>%
                       `colnames<-`(c("Model Term", "VIF(Model Term)")) %>%
                       mutate(Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                       select(Model, `Model Term`, `VIF(Model Term)`) )
  

  #############
  ##  UCH-L1  ##
  #############
  ADbiomarker<-"UCH-L1"
  offset<-0.00001 # Preliminary offset value for UCH-L1 as defined in earlier code chunk.
  formula_adjCytokine_i<-as.formula( gsub("placeholder.AD_Biomarker", ADbiomarker, formula_adjCytokine) )

  # 'AdjCytokine' model + Cytokine:
  lmAdjCytokine.UCHL1[[i]]<-try( lm( formula_adjCytokine_i, data=df[[i]] ) )
  tidyAdjCytokine.UCHL1[[i]]<-try( lmAdjCytokine.UCHL1[[i]] %>%
                             tidy() %>%
                             mutate(AD_Biomarker = ADbiomarker,
                                    Cytokine = cytokine_i[[i]],
                                    Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                             select(AD_Biomarker, Cytokine, Model, everything()) )
  vifAdjCytokine.UCHL1[[i]]<-try( lmAdjCytokine.UCHL1[[i]] %>%
                       car::vif() %>%
                       as.data.frame() %>%
                       rownames_to_column("Model Term") %>%
                       `colnames<-`(c("Model Term", "VIF(Model Term)")) %>%
                       mutate(Model = paste(as.character(formula_adjCytokine_i)[2:3], collapse=" ~ ")) %>%
                       select(Model, `Model Term`, `VIF(Model Term)`) )
  
}
```

###### Extract the results from step 3 above
```{r}
tidyAdjCytokine_All.NfL<-tidyAdjCytokine.NfL %>% rbindlist()
tidyAdjCytokine_All.GFAP<-tidyAdjCytokine.GFAP %>% rbindlist()
tidyAdjCytokine_All.TAU<-tidyAdjCytokine.TAU %>% rbindlist()
tidyAdjCytokine_All.UCHL1<-tidyAdjCytokine.UCHL1 %>% rbindlist()

fixedAdjCytokine_All.NfL<-tidyAdjCytokine_All.NfL %>%
  filter(#effect=="fixed" &
    term!="(Intercept)") %>%
  #select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "P")) #, "df" ))
fixedAdjCytokine_All.GFAP<-tidyAdjCytokine_All.GFAP %>%
  filter(#effect=="fixed" &
    term!="(Intercept)") %>%
  #select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "P")) #, "df" ))
fixedAdjCytokine_All.TAU<-tidyAdjCytokine_All.TAU %>%
  filter(#effect=="fixed" &
    term!="(Intercept)") %>%
  #select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "P")) #, "df" ))
fixedAdjCytokine_All.UCHL1<-tidyAdjCytokine_All.UCHL1 %>%
  filter(#effect=="fixed" &
    term!="(Intercept)") %>%
  #select(-c(effect, group)) %>%
  `colnames<-`(c("AD_Biomarker", "Cytokine", "Model", "Term", "Estimate", "SE", "Statistic", "P")) #, "df" ))

fixedAdjCytokine_All.NfL
fixedAdjCytokine_All.GFAP
fixedAdjCytokine_All.TAU
fixedAdjCytokine_All.UCHL1

cytokineResultsAdjCytokine_All.NfL<-fixedAdjCytokine_All.NfL %>%
  filter(Term!="T21" & Term!="Age" & Term!="Female") %>%
  arrange(P)

T21resultsAdjCytokine_All.NfL<-fixedAdjCytokine_All.NfL %>%
  filter(Term=="T21") %>%
  rename(Estimate_T21_ADJcytokine = Estimate,
         SE_T21_ADJcytokine = SE,
         Statistic_T21_ADJcytokine = Statistic,
         #df_T21_ADJcytokine = df,
         P_T21_ADJcytokine = P) %>%
  arrange(desc(P_T21_ADJcytokine))
T21resultsAdjCytokine_All.GFAP<-fixedAdjCytokine_All.GFAP %>%
  filter(Term=="T21") %>%
  rename(Estimate_T21_ADJcytokine = Estimate,
         SE_T21_ADJcytokine = SE,
         Statistic_T21_ADJcytokine = Statistic,
         #df_T21_ADJcytokine = df,
         P_T21_ADJcytokine = P) %>%
  arrange(desc(P_T21_ADJcytokine))
T21resultsAdjCytokine_All.TAU<-fixedAdjCytokine_All.TAU %>%
  filter(Term=="T21") %>%
  rename(Estimate_T21_ADJcytokine = Estimate,
         SE_T21_ADJcytokine = SE,
         Statistic_T21_ADJcytokine = Statistic,
         #df_T21_ADJcytokine = df,
         P_T21_ADJcytokine = P) %>%
  arrange(desc(P_T21_ADJcytokine))
T21resultsAdjCytokine_All.UCHL1<-fixedAdjCytokine_All.UCHL1 %>%
  filter(Term=="T21") %>%
  rename(Estimate_T21_ADJcytokine = Estimate,
         SE_T21_ADJcytokine = SE,
         Statistic_T21_ADJcytokine = Statistic,
         #df_T21_ADJcytokine = df,
         P_T21_ADJcytokine = P) %>%
  arrange(desc(P_T21_ADJcytokine))

T21resultsAdjCytokine_All.NfL %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
T21resultsAdjCytokine_All.GFAP %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
T21resultsAdjCytokine_All.TAU %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
T21resultsAdjCytokine_All.UCHL1 %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)


cytokineResultsAdjT21_All.NfL<-fixedAdjCytokine_All.NfL %>%
  filter(Term!="T21" & Term!="Age" & Term!="Female") %>%
  rename(Estimate_Cytokine_AdjT21 = Estimate,
         SE_Cytokine_AdjT21 = SE,
         Statistic_Cytokine_AdjT21 = Statistic,
         #df_T21_ADJcytokine = df,
         P_Cytokine_AdjT21 = P) %>%
  arrange(desc(P_Cytokine_AdjT21))

cytokineResultsAdjT21_All.GFAP<-fixedAdjCytokine_All.GFAP %>%
  filter(Term!="T21" & Term!="Age" & Term!="Female") %>%
  rename(Estimate_Cytokine_AdjT21 = Estimate,
         SE_Cytokine_AdjT21 = SE,
         Statistic_Cytokine_AdjT21 = Statistic,
         #df_T21_ADJcytokine = df,
         P_Cytokine_AdjT21 = P) %>%
  arrange(desc(P_Cytokine_AdjT21))

cytokineResultsAdjT21_All.TAU<-fixedAdjCytokine_All.TAU %>%
  filter(Term!="T21" & Term!="Age" & Term!="Female") %>%
  rename(Estimate_Cytokine_AdjT21 = Estimate,
         SE_Cytokine_AdjT21 = SE,
         Statistic_Cytokine_AdjT21 = Statistic,
         #df_T21_ADJcytokine = df,
         P_Cytokine_AdjT21 = P) %>%
  arrange(desc(P_Cytokine_AdjT21))

cytokineResultsAdjT21_All.UCHL1<-fixedAdjCytokine_All.UCHL1 %>%
  filter(Term!="T21" & Term!="Age" & Term!="Female") %>%
  rename(Estimate_Cytokine_AdjT21 = Estimate,
         SE_Cytokine_AdjT21 = SE,
         Statistic_Cytokine_AdjT21 = Statistic,
         #df_T21_ADJcytokine = df,
         P_Cytokine_AdjT21 = P) %>%
  arrange(desc(P_Cytokine_AdjT21))
```

##### Mediation Analysis | Step 4: Identify mediatiors by evaluating changes in results for T21 after adjustment for cytokines
##### NfL
```{R}
temp.ADbiomarkerOnly<-T21resultsBasic_BASELINE.NfL %>%
  select(AD_Biomarker, Model, Estimate_T21, P_T21) %>%
  rename(Model_ADbiomarkerOnly = Model,
         Estimate_T21_ADbiomarkerOnly = Estimate_T21,
         P_T21_ADbiomarkerOnly = P_T21) %>%
  mutate(Sig_ADbiomarkerOnly = ifelse(P_T21_ADbiomarkerOnly<0.05, 1, 0))
temp.CytokineOnly<-T21results.Cytokine %>%
  select(Cytokine, Model, Estimate_T21, P_T21, Sig_Cytokine) %>%
  rename(Model_CytokineOnly = Model,
         Estimate_T21_CytokineOnly = Estimate_T21,
         P_T21_CytokineOnly = P_T21,
         Sig_CytokineOnly = Sig_Cytokine)
tempT21.ADbiomarkerADJcytokine<-T21resultsAdjCytokine_All.NfL %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
tempCytokine.ADbiomarkerAdjT21<-cytokineResultsAdjT21_All.NfL %>% select(AD_Biomarker, Cytokine, Model, Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21)
temp.ADbiomarkerADJcytokine<-tempT21.ADbiomarkerADJcytokine %>%
  full_join(tempCytokine.ADbiomarkerAdjT21, by=c("AD_Biomarker", "Cytokine", "Model")) %>%
  mutate(Sig_T21adjCytokine = ifelse(P_T21_ADJcytokine<0.05, 1, 0),
         Sig_CytokineAdjT21 = ifelse(P_Cytokine_AdjT21<0.05, 1, 0)) %>%
  rename(Mediation_Model = Model) %>%
  select(AD_Biomarker, Cytokine, Mediation_Model,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21)

temp.CytokineOnly
temp.ADbiomarkerOnly
temp.ADbiomarkerADJcytokine

change_T21Results.NfL<-temp.ADbiomarkerADJcytokine %>%
  full_join(temp.ADbiomarkerOnly, by=c("AD_Biomarker")) %>%
  full_join(temp.CytokineOnly, by=c("Cytokine")) %>%
  select(AD_Biomarker, Cytokine,
         Model_CytokineOnly, Estimate_T21_CytokineOnly, P_T21_CytokineOnly, Sig_CytokineOnly,
         Model_ADbiomarkerOnly, Estimate_T21_ADbiomarkerOnly, P_T21_ADbiomarkerOnly, Sig_ADbiomarkerOnly,
         Mediation_Model,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine)
change_T21Results.NfL

change_T21Results.NfL %>%
  filter(Sig_ADbiomarkerOnly==1 &
         Sig_CytokineOnly==1) %>%
  filter(Sig_T21adjCytokine==0 &
           Sig_CytokineAdjT21==1)
```

##### GFAP
```{R}
temp.ADbiomarkerOnly<-T21resultsBasic_BASELINE.GFAP %>%
  select(AD_Biomarker, Model, Estimate_T21, P_T21) %>%
  rename(Model_ADbiomarkerOnly = Model,
         Estimate_T21_ADbiomarkerOnly = Estimate_T21,
         P_T21_ADbiomarkerOnly = P_T21) %>%
  mutate(Sig_ADbiomarkerOnly = ifelse(P_T21_ADbiomarkerOnly<0.05, 1, 0))
temp.CytokineOnly<-T21results.Cytokine %>%
  select(Cytokine, Model, Estimate_T21, P_T21, Sig_Cytokine) %>%
  rename(Model_CytokineOnly = Model,
         Estimate_T21_CytokineOnly = Estimate_T21,
         P_T21_CytokineOnly = P_T21,
         Sig_CytokineOnly = Sig_Cytokine)
tempT21.ADbiomarkerADJcytokine<-T21resultsAdjCytokine_All.GFAP %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
tempCytokine.ADbiomarkerAdjT21<-cytokineResultsAdjT21_All.GFAP %>% select(AD_Biomarker, Cytokine, Model, Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21)
temp.ADbiomarkerADJcytokine<-tempT21.ADbiomarkerADJcytokine %>%
  full_join(tempCytokine.ADbiomarkerAdjT21, by=c("AD_Biomarker", "Cytokine", "Model")) %>%
  mutate(Sig_T21adjCytokine = ifelse(P_T21_ADJcytokine<0.05, 1, 0),
         Sig_CytokineAdjT21 = ifelse(P_Cytokine_AdjT21<0.05, 1, 0)) %>%
  rename(Mediation_Model = Model) %>%
  select(AD_Biomarker, Cytokine, Mediation_Model,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21)

temp.CytokineOnly
temp.ADbiomarkerOnly
temp.ADbiomarkerADJcytokine

### PICK UP HERE (8/12/20 ~6PM): Join together the different model results, then identify mediators:
change_T21Results.GFAP<-temp.ADbiomarkerADJcytokine %>%
  full_join(temp.ADbiomarkerOnly, by=c("AD_Biomarker")) %>%
  full_join(temp.CytokineOnly, by=c("Cytokine")) %>%
  select(AD_Biomarker, Cytokine,
         Model_CytokineOnly, Estimate_T21_CytokineOnly, P_T21_CytokineOnly, Sig_CytokineOnly,
         Model_ADbiomarkerOnly, Estimate_T21_ADbiomarkerOnly, P_T21_ADbiomarkerOnly, Sig_ADbiomarkerOnly,
         Mediation_Model,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine)
change_T21Results.GFAP

change_T21Results.GFAP %>%
  filter(Sig_ADbiomarkerOnly==1 &
         Sig_CytokineOnly==1) %>%
  filter(Sig_T21adjCytokine==0 &
           Sig_CytokineAdjT21==1)
```

##### TAU
```{R}
temp.ADbiomarkerOnly<-T21resultsBasic_BASELINE.TAU %>%
  select(AD_Biomarker, Model, Estimate_T21, P_T21) %>%
  rename(Model_ADbiomarkerOnly = Model,
         Estimate_T21_ADbiomarkerOnly = Estimate_T21,
         P_T21_ADbiomarkerOnly = P_T21) %>%
  mutate(Sig_ADbiomarkerOnly = ifelse(P_T21_ADbiomarkerOnly<0.05, 1, 0))
temp.CytokineOnly<-T21results.Cytokine %>%
  select(Cytokine, Model, Estimate_T21, P_T21, Sig_Cytokine) %>%
  rename(Model_CytokineOnly = Model,
         Estimate_T21_CytokineOnly = Estimate_T21,
         P_T21_CytokineOnly = P_T21,
         Sig_CytokineOnly = Sig_Cytokine)
tempT21.ADbiomarkerADJcytokine<-T21resultsAdjCytokine_All.TAU %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
tempCytokine.ADbiomarkerAdjT21<-cytokineResultsAdjT21_All.TAU %>% select(AD_Biomarker, Cytokine, Model, Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21)
temp.ADbiomarkerADJcytokine<-tempT21.ADbiomarkerADJcytokine %>%
  full_join(tempCytokine.ADbiomarkerAdjT21, by=c("AD_Biomarker", "Cytokine", "Model")) %>%
  mutate(Sig_T21adjCytokine = ifelse(P_T21_ADJcytokine<0.05, 1, 0),
         Sig_CytokineAdjT21 = ifelse(P_Cytokine_AdjT21<0.05, 1, 0)) %>%
  rename(Mediation_Model = Model) %>%
  select(AD_Biomarker, Cytokine, Mediation_Model,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21)

temp.CytokineOnly
temp.ADbiomarkerOnly
temp.ADbiomarkerADJcytokine

change_T21Results.TAU<-temp.ADbiomarkerADJcytokine %>%
  full_join(temp.ADbiomarkerOnly, by=c("AD_Biomarker")) %>%
  full_join(temp.CytokineOnly, by=c("Cytokine")) %>%
  select(AD_Biomarker, Cytokine,
         Model_CytokineOnly, Estimate_T21_CytokineOnly, P_T21_CytokineOnly, Sig_CytokineOnly,
         Model_ADbiomarkerOnly, Estimate_T21_ADbiomarkerOnly, P_T21_ADbiomarkerOnly, Sig_ADbiomarkerOnly,
         Mediation_Model,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine)
change_T21Results.TAU

change_T21Results.TAU %>%
  filter(Sig_ADbiomarkerOnly==1 &
         Sig_CytokineOnly==1) %>%
  filter(Sig_T21adjCytokine==0 &
           Sig_CytokineAdjT21==1)
```

##### UCH-L1
```{R}
temp.ADbiomarkerOnly<-T21resultsBasic_BASELINE.UCHL1 %>%
  select(AD_Biomarker, Model, Estimate_T21, P_T21) %>%
  rename(Model_ADbiomarkerOnly = Model,
         Estimate_T21_ADbiomarkerOnly = Estimate_T21,
         P_T21_ADbiomarkerOnly = P_T21) %>%
  mutate(Sig_ADbiomarkerOnly = ifelse(P_T21_ADbiomarkerOnly<0.05, 1, 0))
temp.CytokineOnly<-T21results.Cytokine %>%
  select(Cytokine, Model, Estimate_T21, P_T21, Sig_Cytokine) %>%
  rename(Model_CytokineOnly = Model,
         Estimate_T21_CytokineOnly = Estimate_T21,
         P_T21_CytokineOnly = P_T21,
         Sig_CytokineOnly = Sig_Cytokine)
tempT21.ADbiomarkerADJcytokine<-T21resultsAdjCytokine_All.UCHL1 %>% select(AD_Biomarker, Cytokine, Model, Estimate_T21_ADJcytokine, P_T21_ADJcytokine)
tempCytokine.ADbiomarkerAdjT21<-cytokineResultsAdjT21_All.UCHL1 %>% select(AD_Biomarker, Cytokine, Model, Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21)
temp.ADbiomarkerADJcytokine<-tempT21.ADbiomarkerADJcytokine %>%
  full_join(tempCytokine.ADbiomarkerAdjT21, by=c("AD_Biomarker", "Cytokine", "Model")) %>%
  mutate(Sig_T21adjCytokine = ifelse(P_T21_ADJcytokine<0.05, 1, 0),
         Sig_CytokineAdjT21 = ifelse(P_Cytokine_AdjT21<0.05, 1, 0)) %>%
  rename(Mediation_Model = Model) %>%
  select(AD_Biomarker, Cytokine, Mediation_Model,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21)

temp.CytokineOnly
temp.ADbiomarkerOnly
temp.ADbiomarkerADJcytokine

### PICK UP HERE (8/12/20 ~6PM): Join together the different model results, then identify mediators:
change_T21Results.UCHL1<-temp.ADbiomarkerADJcytokine %>%
  full_join(temp.ADbiomarkerOnly, by=c("AD_Biomarker")) %>%
  full_join(temp.CytokineOnly, by=c("Cytokine")) %>%
  select(AD_Biomarker, Cytokine,
         Model_CytokineOnly, Estimate_T21_CytokineOnly, P_T21_CytokineOnly, Sig_CytokineOnly,
         Model_ADbiomarkerOnly, Estimate_T21_ADbiomarkerOnly, P_T21_ADbiomarkerOnly, Sig_ADbiomarkerOnly,
         Mediation_Model,
         Estimate_Cytokine_AdjT21, P_Cytokine_AdjT21, Sig_CytokineAdjT21,
         Estimate_T21_ADJcytokine, P_T21_ADJcytokine, Sig_T21adjCytokine)
change_T21Results.UCHL1

change_T21Results.UCHL1 %>%
  filter(Sig_ADbiomarkerOnly==1 &
         Sig_CytokineOnly==1) %>%
  filter(Sig_T21adjCytokine==0 &
           Sig_CytokineAdjT21==1)
```
