---
title: "R Notebook"
output: html_notebook
---

```{r}
#library(limma)
library(dplyr)
library(tidyr)
library(data.table)
#install.packages(c("broom", "broomExtra", "tibble", "sjstats", "car", "lme4", "lmerTest"))
library(broom)
#install.packages("broomExtra")
#library(broomExtra)
library(tibble)
library(sjstats)
library(car)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tibble)
library(modelr)
#install.packages("tidyverse")
#library(tidyverse)
#library(miceadds)
#library(ggforce)
```

```{r}
help("vif") # car
help(AIC) # stats
```

```{r}
#### Read in dataset of sample Source (for batch effect correction):
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/Demographics")
sample_source<-fread("LAB_ID Source 4.30.20.csv") %>%
  rename(RecordID=`Record ID`, LabID=`Lab ID`, Date=`Date Obtained`, Source=`Source vlookup`) %>%
  mutate(RecordID=gsub("NDAR_", "", RecordID)) %>%
  mutate(I_SourceData=1)

#### Read in master meta-data:
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/META/Re__MASTER_RecordID_vs_LabID.Labels.05.11.20")
meta<-fread("HTP_CLEANED_051120_v0.3_MASTER_RecordID_vs_LabID.Labels.csv") %>%
  rename(Age=Age_at_visit, Date=Date_obtained) %>%
  mutate(T21=ifelse(Cohort_type=="Down syndrome", 1,
                    ifelse(Cohort_type=="Control", 0, NA)),
         Female=ifelse(Sex=="Female", 1,
                       ifelse(Sex=="Male", 0, NA))) %>%
  select(-c(Cohort_type)) %>%
  mutate(I_MetaData=1)

#### Read in MSD data and join with master meta-data and sample Source dataset:
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/MSD/Derived/Master/200413/Clean/")
msd<-fread("200413_CLEAN_noMeta_Exp3456_firstPIDsampleOnly_meansOnly.csv") %>%
  rename(RecordID=PID) %>%
  select(-T21) %>%
  filter(Panel_Analyte!="Angiogenesis_VEGF-A") %>% # VEGF-A was included on two different panels. Both have excellent detection rates, but the VEGF-A from the Cytokine 1 panel has perfect detection. So, I'm keeping Cytokine 1_VEGF-A and removing Angiogenesis_VEGF-A.
  mutate(I_MSDdata=1) %>%
  left_join(meta, by=c("RecordID", "LabID")) %>%
  left_join(sample_source, by=c("RecordID", "LabID", "Date")) %>%
  filter(Source=="Local" | Source=="NDSC 2018" | Source=="NDSC 2019") # Keep only samples obtained locally or at NDSC

msd

by_analyte<-msd %>%
  group_by(Panel_Analyte)
```

```{r}
# setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/MSD/Derived/Master/200413/Clean/")
# msd<-fread("200413_CLEAN_noMeta_Exp3456_firstPIDsampleOnly_meansOnly.csv")
# msd
```

```{r}
# samp1.MeansOnly<-dat %>%
#   filter(Panel_Analyte!="Chemokine_IL-8" &
#            Panel_Analyte!="TH17_IL-17A" &
#            Panel_Analyte!="Cytokine2_IL-17A/F" &
#            Analyte!="IL-21") %>%
#   filter(Source=="Local" | Source=="NDSC 2018" | Source=="NDSC 2019") %>%
#   rename(Age=Age)
# 
# samp1.MeansOnly %>% dplyr::select(Panel, Analyte) %>% unique() # 55 analytes.
# 
# samp1.MeansOnly
# 
# by_analyte<-samp1.MeansOnly %>%
#   group_by(Panel_Analyte)
```

# Identify the best model parameterization using AIC

# Summarize changes in fit statistics with the addition of random intercept for source (denoted '+ (1|Source)')
```{r}
set.seed(1234)

m1 = by_analyte %>%
  do(fit = lm(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female, data = .))
names(m1$fit)<-m1$Panel_Analyte
call1<-(m1$fit[[1]]$call %>% as.character())[2]

m2 = by_analyte %>%
  do(fit = lmerTest::lmer(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + (1|Source), REML=FALSE, data = .))
names(m2$fit)<-m2$Panel_Analyte
call2<-(summary(m2$fit[[1]])$call %>% as.character())[2]

temp.AIC<-vector("list", 54)
temp.maxVIF1<-vector("list", 54)
temp.maxVIF2<-vector("list", 54)
temp.Rsq1<-vector("list", 54)
temp.Rsq2<-vector("list", 54)
temp.sigma<-vector("list", 54)
for ( i in 1:length(m1$fit) ) {
  temp.AIC[[i]]<-AIC( m1$fit[[i]],
                         m2$fit[[i]]) %>%
    rownames_to_column() %>%
    rename(Model=rowname) %>%
    mutate(Model_Call = ifelse(Model=="m1$fit[[i]]", call1,
                               ifelse(Model=="m2$fit[[i]]", call2, NA))) %>%
    mutate(Panel_Analyte = m1$Panel_Analyte[i]) %>%
    select(Panel_Analyte, Model, Model_Call, df, AIC)
  compare.AIC<-rbindlist(temp.AIC)
  
  temp.maxVIF1[[i]]<-max(vif(m1$fit[[i]])) %>%
    as.data.frame() %>%
    `colnames<-`("Max_VIF") %>%
    mutate(Model = "m1$fit[[i]]", Panel_Analyte = m1$Panel_Analyte[i]) %>%
    select(Panel_Analyte, Model, Max_VIF)
  temp.maxVIF2[[i]]<-max(vif(m2$fit[[i]])) %>%
    as.data.frame() %>%
    `colnames<-`("Max_VIF") %>%
    mutate(Model = "m2$fit[[i]]", Panel_Analyte = m2$Panel_Analyte[i]) %>%
    select(Panel_Analyte, Model, Max_VIF)
  compare.maxVIF<-rbind( rbindlist(temp.maxVIF1),
                  rbindlist(temp.maxVIF2) )

  temp.Rsq1[[i]]<-cbind("m1", performance::r2(m1$fit[[i]])$R2, performance::r2(m1$fit[[i]])$R2_adjusted) %>%
    as.data.frame() %>%
    `colnames<-`(c("Model", "R2", "R2_adjusted")) %>%
    mutate(Panel_Analyte = m1$Panel_Analyte[i]) %>%
    select(Panel_Analyte, Model, R2, R2_adjusted)
  temp.Rsq2[[i]]<-cbind("m2", performance::r2(m2$fit[[i]])$R2_conditional, performance::r2(m2$fit[[i]])$R2_marginal) %>%
    as.data.frame() %>%
    `colnames<-`(c("Model", "R2_conditional", "R2_marginal")) %>%
    mutate(Panel_Analyte = m2$Panel_Analyte[i]) %>%
    select(Panel_Analyte, Model, R2_conditional, R2_marginal)
  m1.R2<-rbindlist(temp.Rsq1)
  m2.R2<-rbindlist(temp.Rsq2)
  
  temp.sigma[[i]]<-cbind(m1$Panel_Analyte[i],
                         m2$Panel_Analyte[i],
                         summary(m1$fit[[i]])$sigma,
                         summary(m2$fit[[i]])$sigma) %>%
    `colnames<-`(c("Panel_Analyte", "verify_Panel_Analyte", "sigma1", "sigma2")) %>%
    as.data.frame() %>%
    mutate(sigma1=as.numeric(as.character(sigma1)),
           sigma2=as.numeric(as.character(sigma2)),
           sigma2_minus_sigma1 = sigma2 - sigma1,
           pctChange_sigma = 100*sigma2_minus_sigma1 / sigma1)
           
  compare.sigma<-rbindlist(temp.sigma)
}

compare.AIC_wide<-compare.AIC %>% select(Panel_Analyte, Model, AIC) %>% unique() %>%
  spread(key="Model", value="AIC") %>%
  mutate(AIC2_minus_AIC1=`m2$fit[[i]]`-`m1$fit[[i]]`)
compare.AIC_wide

compare.sigma

vif(m1$fit[[1]])
vif(m2$fit[[1]])
# T21      Age   Female 
# 1.060194 1.049741 1.043231 
#     T21      Age   Female 
# 1.038281 1.029492 1.039372

vif(m1$fit[[2]])
vif(m2$fit[[2]])

compare.maxVIF %>%
  mutate(Model = ifelse(Model == "m1$fit[[i]]", "1",
                        ifelse(Model == "m2$fit[[i]]", "2", NA)))
```


```{r}
compare.AIC_wide01<-compare.AIC_wide %>%
  rename(Model1=`m1$fit[[i]]`, Model2=`m2$fit[[i]]`) %>%
  mutate(`deltaAIC<=-2` = ifelse(AIC2_minus_AIC1<=-2, 1, 0),
         `deltaAIC<=-5` = ifelse(AIC2_minus_AIC1<=-5, 1, 0),
         `deltaAIC<=-10` = ifelse(AIC2_minus_AIC1<=-10, 1, 0),
         `deltaAIC>=2` = ifelse(AIC2_minus_AIC1>=2, 1, 0),
         `deltaAIC>=5` = ifelse(AIC2_minus_AIC1>=5, 1, 0),
         `deltaAIC>=10` = ifelse(AIC2_minus_AIC1>=10, 1, 0)) %>%
  rename(`deltaAIC (AIC2 - AIC1)` = AIC2_minus_AIC1)

compare.deltaAIC2<-compare.AIC_wide01 %>%
  group_by(`deltaAIC<=-2`, `deltaAIC>=2`) %>%
  summarise(N=n()) %>%
  mutate(Category = ifelse(`deltaAIC<=-2`==0 & `deltaAIC>=2`==0, "Ambivalent",
                           ifelse(`deltaAIC<=-2`==1, "Better fit",
                                  ifelse(`deltaAIC>=2`==1, "Worse fit", NA))))

compare.deltaAIC5<-compare.AIC_wide01 %>%
  group_by(`deltaAIC<=-5`, `deltaAIC>=5`) %>%
  summarise(N=n()) %>%
  mutate(Category = ifelse(`deltaAIC<=-5`==0 & `deltaAIC>=5`==0, "Ambivalent",
                           ifelse(`deltaAIC<=-5`==1, "Better fit",
                                  ifelse(`deltaAIC>=5`==1, "Worse fit", NA))))

compare.deltaAIC10<-compare.AIC_wide01 %>%
  group_by(`deltaAIC<=-10`, `deltaAIC>=10`) %>%
  summarise(N=n()) %>%
  mutate(Category = ifelse(`deltaAIC<=-10`==0 & `deltaAIC>=10`==0, "Ambivalent",
                           ifelse(`deltaAIC<=-10`==1, "Better fit",
                                  ifelse(`deltaAIC>=10`==1, "Worse fit", NA))))

compare.AIC_wide01
compare.deltaAIC2
compare.deltaAIC5
compare.deltaAIC10
```







```{r}
lm4 = by_analyte %>%
  do(fit3 = lm(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + T21*Age, data = .))

lmm = by_analyte %>%
  do(fit4 = lme4::lmer(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + T21*Age + (1|Source), data = ., REML=FALSE))

AIC1 =  broom::glance(lm1, fit1) %>% select(Panel_Analyte, AIC) %>% rename(AIC1=AIC)
AIC2 =  broom::glance(lm2, fit2) %>% select(Panel_Analyte, AIC) %>% rename(AIC2=AIC)
AIC3 =  broom::glance(lm3, fit3) %>% select(Panel_Analyte, AIC) %>% rename(AIC3=AIC)
AIC4 =  broom::glance(lmm, fit4) %>% select(Panel_Analyte, AIC) %>% rename(AIC4=AIC)

AIC<-AIC1 %>%
  full_join(AIC2, by="Panel_Analyte") %>%
  full_join(AIC3, by="Panel_Analyte") %>%
  full_join(AIC4, by="Panel_Analyte") %>%
  mutate(BetterWithFemale=ifelse(AIC2<AIC1, 1, 0),
         SigBetterWithFemale=ifelse( (AIC2-AIC1)<=-2, 1, 0)) %>%
  mutate(WorseWithFemale=ifelse(AIC2>AIC1, 1, 0),
         SigWorseWithFemale=ifelse( (AIC2-AIC1)>=2, 1, 0)) %>%
  mutate(SigBetterWithInteraction=ifelse( (AIC3-AIC2)<=-2, 1, 0),
         SigWorseWithInteraction=ifelse( (AIC3-AIC2)>=2, 1, 0)) %>%
  mutate(SigBetterWithSource=ifelse( (AIC4-AIC3)<=-2, 1, 0),
         SigWorseWithSource=ifelse( (AIC4-AIC3)>=2, 1, 0))
AIC %>% group_by(SigBetterWithFemale) %>% summarise(N=n()) # 13 significantly better after including Female covariate
AIC %>% group_by(SigWorseWithFemale) %>% summarise(N=n()) # 0 significantly worse after including Female covariate

AIC %>% group_by(SigBetterWithInteraction) %>% summarise(N=n()) # 9 significantly better after including T21*Age
AIC %>% group_by(SigWorseWithInteraction) %>% summarise(N=n()) # 0 significantly worse after including T21*Age

AIC %>% group_by(SigBetterWithSource) %>% summarise(N=n()) # 40 significantly better after including (1|Source)
AIC %>% group_by(SigWorseWithSource) %>% summarise(N=n()) # 4 significantly worse after including (1|Source)
AIC %>% filter(SigWorseWithSource==1) %>% select(Panel_Analyte)
# Significantly worse by AIC after inclusion of random intercept for Source:
# Cytokine 1_IL-12/IL-23p40				
# TH17_IL-22				
# TH17_MIP-3alpha				
# VascularInjury_ICAM-1				

AIC %>% filter(SigBetterWithSource==0 & SigWorseWithSource==0) %>% select(Panel_Analyte)
# Cytokines that neither improve nor worsen with (1|Source):
# Angiogenesis_VEGF-D				
# Chemokine_Eotaxin				
# Cytokine 1_IL-15				
# Cytokine 1_TNF-beta				
# Cytokine 1_VEGF-A				
# Proinflammatory_IL-10				
# Proinflammatory_IL-1beta				
# Proinflammatory_IL-2				
# Proinflammatory_IL-6				
# Proinflammatory_TNF-alpha

# Chosen model: log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + T21*Age + (1|Source)
```

#### Confirm that results are the same with ML and REML
```{r}
lmm_ML = by_analyte %>%
  do(fit.lmm_ML = lmerTest::lmer(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + T21*Age + (1|Source), data = ., REML=FALSE))

lmm_REML = by_analyte %>%
  do(fit.lmm_REML = lmerTest::lmer(log2(CalcConcMean_AfterImputation) ~ T21 + Age + Female + T21*Age + (1|Source), data = ., REML=TRUE))

# augment:
augment.lmm_ML = augment(lmm_ML, fit.lmm_ML)
augment.lmm_REML = augment(lmm_REML, fit.lmm_REML)

# glance:
glance.lmm_ML = broom::glance(lmm_ML, fit.lmm_ML)
glance.lmm_REML = broom::glance(lmm_REML, fit.lmm_REML)

# Name each dataframe in list of results:
names(lmm_ML$fit.lmm_ML)<-lmm_ML$Panel_Analyte
names(lmm_REML$fit.lmm_REML)<-lmm_REML$Panel_Analyte

# tidy:
# Define new function that defaults to include confidence intervals in output:
tidyPlus <- function(x, print=TRUE) {
  result <- tidy(x, conf.int=T)
  return(result)
}
# Apply tidyPlus to all dataframes in list 'dfAnalyte_lm$fitAnalyte_lm'
tidy.lmm_ML<-lapply(lmm_ML$fit.lmm_ML, tidyPlus) %>%
  bind_rows(., .id="Panel_Analyte")
tidy.lmm_REML<-lapply(lmm_REML$fit.lmm_REML, tidyPlus) %>%
  bind_rows(., .id="Panel_Analyte")

tidy.lmm_ML
tidy.lmm_REML

T21_results.lmm_ML<-tidy.lmm_ML %>% filter(term=="T21") %>% select(-c(effect, group, statistic, df)) %>%
  rename(P=p.value) %>% mutate(P_BH=p.adjust(P, method="BH")) %>%
  arrange(P_BH) %>%
  mutate(Rank_ML=seq(1:55))
T21_results.lmm_REML<-tidy.lmm_ML %>% filter(term=="T21") %>% select(-c(effect, group, statistic, df)) %>%
  rename(P=p.value) %>% mutate(P_BH=p.adjust(P, method="BH")) %>%
  arrange(P_BH) %>%
  mutate(Rank_REML=seq(1:55))

T21_results.lmm_ML
T21_results.lmm_REML

rank.MLvsREML<-T21_results.lmm_ML %>% select(Panel_Analyte, Rank_ML, estimate) %>%
  rename(Estimate_ML=estimate) %>%
  full_join(T21_results.lmm_REML, by="Panel_Analyte") %>%
  rename(Estimate_REML=estimate) %>%
  select(Panel_Analyte, Rank_ML, Rank_REML, Estimate_ML, Estimate_REML) %>%
  mutate(delta_Estimate=Estimate_REML-Estimate_ML)
rank.MLvsREML # Exactly the same ranks and estimates - Good.
```

#### Look at how many T21s vs. D21s were sampled at each Source location
```{r}
by_analyte %>%
  ungroup() %>%
  select(PID, Source, T21) %>%
  unique() %>%
  group_by(Source, T21) %>%
  summarise(N=n())
```

#### Repeat the model fitting procedure to be sure that the apparent effect of Source isn't due to distribution of T21 vs. D21 within each Source
```{r}
lm_T21 = by_analyte %>%
  filter(T21==1) %>%
  do(fit.lm_T21 = lm(log2(CalcConcMean_AfterImputation) ~ Age + Female, data = .))

lmm_T21 = by_analyte %>%
  filter(T21==1) %>%
  do(fit.lmm_T21 = lme4::lmer(log2(CalcConcMean_AfterImputation) ~ Age + Female + (1|Source), data = ., REML=FALSE))

AIC.lm_T21 =  broom::glance(lm_T21, fit.lm_T21) %>% select(Panel_Analyte, AIC) %>% rename(AIC_lmT21=AIC)
AIC.lmm_T21 =  broom::glance(lmm_T21, fit.lmm_T21) %>% select(Panel_Analyte, AIC) %>% rename(AIC_lmmT21=AIC)

AIC<-AIC.lm_T21 %>%
  full_join(AIC.lmm_T21, by="Panel_Analyte") %>%
  mutate(SigBetterWithSource=ifelse( (AIC_lmmT21-AIC_lmT21)<=-2, 1, 0),
         SigWorseWithSource=ifelse( (AIC_lmmT21-AIC_lmT21)>=2, 1, 0))
AIC %>% group_by(SigBetterWithSource, SigWorseWithSource) %>% summarise(N=n())
# SigBetterWithSource SigWorseWithSource  N
# 0	                  0	                  13		
# 0	                  1	                  3		
# 1	                  0	                  39

# So even considering only T21s, (1|Source) improves model fit for the vast majority of analytes.
# 3 analytes are less happy after adjustment for Source, but the change in AIC is just barely 'significant':
AIC %>% filter(SigWorseWithSource==1)
# Panel_Analyte               AIC_lmT21   AIC_lmmT21  SigBetterWithSource   SigWorseWithSource
# Cytokine 1_IL-12/IL-23p40	  620.0699	  622.0699	  0	                    1
# Cytokine 1_IL-15	          383.4355	  385.4355	  0	                    1
```

#### Final touches
```{r}
#setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/LMM/Results")
#T21_results.lmm_ML<-fread("200302_T21results_LMM_FIXEDageFemaleT21byFemale_RANDsource.csv")

T21_results.lmm_ML01<-T21_results.lmm_ML %>%
  #rename(Rank=Rank_ML) %>%
  mutate(Sig_BH=ifelse(P_BH<0.05, 1, 0))
T21_results.lmm_ML01
```

#### Now output the results
```{r}
setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/LMM/Results")
fwrite(T21_results.lmm_ML01, "200305_T21results_LMM_FIXEDageFemaleT21byFemale_RANDsource.csv")
```

#### Model diagnostics
```{r}
qqnorm(augment.lm3$`.resid`)
qqline(augment.lm3$`.resid`)

qqnorm(augment.lmm_ML$`.resid`)
qqline(augment.lmm_ML$`.resid`)

# "p.value	 an approximate p-value for the test. This is said in Royston (1995) to be adequate for p.value < 0.1."
shapiro.test(rnorm(100, mean = 5, sd = 3))
shapiro.test(runif(100, min = 2, max = 4))
# Shapiro-Wilk normality test
# data:  rnorm(100, mean = 5, sd = 3)
# W = 0.97525, p-value = 0.0564
#	Shapiro-Wilk normality test
# data:  runif(100, min = 2, max = 4)
# W = 0.92874, p-value = 4.222e-05

class(runif(100, min = 2, max = 4)) #[1] "numeric"

augment.LM<-augment.lm3 %>% rename(resid=`.resid`)
augment.LMM<-augment.lmm_ML %>% rename(resid=`.resid`)
# "p.value	 an approximate p-value for the test. This is said in Royston (1995) to be adequate for p.value < 0.1."
resid_lm<-augment.LM$resid
resid_lmm<-augment.LMM$resid
class(resid_lm) #[1] "numeric"
class(resid_lmm) #[1] "numeric"

analyte.index<-augment.LM %>% ungroup() %>% select(Panel_Analyte) %>% unique() %>% dplyr::mutate(Index=seq(1:55))

augment.LM01<-augment.LM %>% ungroup() %>% full_join(analyte.index, by="Panel_Analyte")
augment.LMM01<-augment.LMM %>% ungroup() %>% full_join(analyte.index, by="Panel_Analyte")

shapiro.LM<-list()
shapiro.LMM<-list()
for (i in 1:55) {
  aug.LM<-augment.LM01 %>% filter(Index==i)
  aug.LMM<-augment.LMM01 %>% filter(Index==i)
  shapiro.LM[[i]]<-shapiro.test(aug.LM$resid) %>% unlist() %>% as.matrix() %>% t() %>% as.data.frame() %>%
    mutate(Panel_Analyte=aug.LM$Panel_Analyte[[1]])
  shapiro.LMM[[i]]<-shapiro.test(aug.LMM$resid) %>% unlist() %>% as.matrix() %>% t() %>% as.data.frame() %>%
    mutate(Panel_Analyte=aug.LM$Panel_Analyte[[1]])
}

results_shapiro.LM<-rbindlist(shapiro.LM) %>%
  mutate(p.value=as.numeric(as.character(p.value))) %>%
  mutate(Sig_Shapiro=ifelse(p.value<0.05, 1, 0))
results_shapiro.LMM<-rbindlist(shapiro.LMM) %>%
  mutate(p.value=as.numeric(as.character(p.value))) %>%
  mutate(Sig_Shapiro=ifelse(p.value<0.05, 1, 0))

results_shapiro.LM %>% group_by(Sig_Shapiro) %>% summarise(N=n())
results_shapiro.LMM %>% group_by(Sig_Shapiro) %>% summarise(N=n())
# Most cytokines fail formal test for normality of residuals. Residual plots do look decent, though...
```

```{r}
# augment:
augment.lmm_ML = augment(lmm_ML, fit.lmm_ML)
augment.lmm_REML = augment(lmm_REML, fit.lmm_REML)

# glance:
glance.lmm_ML = broom::glance(lmm_ML, fit.lmm_ML)
glance.lmm_REML = broom::glance(lmm_REML, fit.lmm_REML)

augment.lmm_ML
augment.lmm_REML
glance.lmm_ML
glance.lmm_REML

plot(augment.lmm_ML$`.resid`)
plot(augment.lmm_REML$`.resid`)

plot(density(augment.lmm_ML$`.resid`))
plot(density(augment.lmm_REML$`.resid`))

augment.lm3 = augment(lm3, fit3)
plot(augment.lm3$`.resid`)
plot(density(augment.lm3$`.resid`))

qqnorm(augment.lm3$`.resid`)
qqline(augment.lm3$`.resid`)
```


 