---
title: "R Notebook"
output: html_notebook
---

```{r}
#rm(list = ls(all = TRUE))
library(openxlsx)
library(data.table)
library(dplyr)
library(tidyr)
#library(lattice)
library(ggplot2)
#library(infer)
library(miceadds)
require(dplyr)
library(FedData)
library(ggrepel)
# https://cran.r-project.org/web/packages/sjPlot/vignettes/plot_interactions.html
#library(sjPlot)
library(sjmisc)
library(ggplot2)
```

# Read in clean Crnic MSD data from experiments 3-6
```{r}
#### Read in dataset of sample Source (for batch effect correction):
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/Demographics")
sample_source<-fread("LAB_ID Source 4.30.20.csv") %>%
  rename(RecordID=`Record ID`, LabID=`Lab ID`, Date=`Date Obtained`, Source=`Source vlookup`) %>%
  mutate(RecordID=gsub("NDAR_", "", RecordID)) %>%
  mutate(I_SourceData=1)

#### Read in master meta-data:
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/META/Re__MASTER_RecordID_vs_LabID.Labels.05.11.20")
meta<-fread("HTP_CLEANED_051120_v0.3_MASTER_RecordID_vs_LabID.Labels.csv") %>%
  rename(Age=Age_at_visit, Date=Date_obtained) %>%
  mutate(T21=ifelse(Cohort_type=="Down syndrome", 1,
                    ifelse(Cohort_type=="Control", 0, NA)),
         Female=ifelse(Sex=="Female", 1,
                       ifelse(Sex=="Male", 0, NA))) %>%
  select(-c(Cohort_type)) %>%
  mutate(I_MetaData=1)

#### Read in MSD data and join with master meta-data and sample Source dataset:
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/MSD/Derived/Master/200413/Clean/")
msd<-fread("200413_CLEAN_noMeta_Exp3456_firstPIDsampleOnly_meansOnly.csv") %>%
  rename(RecordID=PID) %>%
  select(-T21) %>%
  filter(Panel_Analyte!="Angiogenesis_VEGF-A") %>% # VEGF-A was included on two different panels. Both have excellent detection rates, but the VEGF-A from the Cytokine 1 panel has perfect detection. So, I'm keeping Cytokine 1_VEGF-A and removing Angiogenesis_VEGF-A.
  mutate(I_MSDdata=1) %>%
  left_join(meta, by=c("RecordID", "LabID")) %>%
  left_join(sample_source, by=c("RecordID", "LabID", "Date")) %>%
  filter(Source=="Local" | Source=="NDSC 2018" | Source=="NDSC 2019") # Keep only samples obtained locally or at NDSC

msd
```

# Prepare data to be plotted, including log2 fold changes, 95% confidence intervals, and corrected P-values (BH)
```{r}
setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200512/Results")
resultsT21.model2<-read.xlsx("MSD_200512_RESULTS_SUMMARY_Models1231A1B_LMM_JRS.xlsx", sheet="Model 2") %>%
  select(Panel_Analyte,
         log2FC_T21, CI.low_T21, CI.high_T21, FDR_T21,
         log2FC_Age, CI.low_Age, CI.high_Age, FDR_Age,
         `log2FC_T21:Age`, `CI.low_T21:Age`, `CI.high_T21:Age`, `FDR_T21:Age`) %>%
  unique() %>%
  filter(`FDR_T21:Age`<0.05) %>%
  arrange(`FDR_T21:Age`) %>%
  mutate(Rank_T21byAge = seq(1:length(`FDR_T21:Age`))) %>%
  # mutate(`log2FC (95% CI)` = paste( round(`log2FC_T21:Age`, digits=2),
  #                                   " (", round(CI.low_T21, digits=2), ", ",
  #                                   round(CI.high_T21, digits=2), ")", sep="")) %>%
  mutate(FDR_T21 = ifelse(FDR_T21<0.00001, formatC(FDR_T21, format = "e", digits = 2), round(FDR_T21, digits=4)),
         FDR_Age = ifelse(FDR_T21<0.00001, formatC(FDR_Age, format = "e", digits = 2), round(FDR_Age, digits=4)),
         `FDR_T21:Age` = ifelse(FDR_T21<0.00001, formatC(`FDR_T21:Age`, format = "e", digits = 2), round(`FDR_T21:Age`, digits=4))) %>%
  separate(Panel_Analyte, into=c("Panel", "Analyte"), sep="_", extra="merge", remove=TRUE)
resultsT21.model2

msd01<-resultsT21.model2 %>%
  left_join(msd, by=c("Panel", "Analyte"))
msd01
```

# Define where to output plots
```{r}
out.dir<-"/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200512/Plots/Age_trajectories"
```

# Define colors for plotting
```{r}
# https://cran.r-project.org/web/packages/broom.mixed/broom.mixed.pdf
# .fitted = predicted values 
# .fixed = predicted values with no random effects

#fills<-c("snow3", "snow3")
fills<-c("gray42", "#007db8")
values<-c("gray42", "#007db8")
colors<-c("#1874CD", "#007db8")
RedBlue<-c("#CD3333", "#1874CD")
Purple<-"#551A8B"
GrayBlue<-c("gray42", "#1874CD")
```

```{r}
# Analytes with significant T21*Age interactions:
# VascularInjury_SAA
# UPLEX_IL-29
# TH17_IL-22
# TH17_IL-27
# Angiogenesis_VEGF-C
```

# Plot age-associated trajectories
```{r}
theme_set(theme_gray(base_size = 12, base_family = "Arial") +
            theme(panel.border = element_rect(colour="black", fill = "transparent"), 
                  plot.title = element_text(face="bold", hjust = 0), # lineheight=.8, size=20,
                  axis.text = element_text(color="black", size = 14), 
                  axis.text.x = element_text(angle = 0, hjust = 0.5),
                  panel.background = element_blank(),
                  panel.grid = element_blank(),
                  plot.background = element_blank()
            ) +
            theme(strip.background = element_rect(colour="black", fill = "light grey", size = 1)) # adjusts facet label borders
)

set.seed(1234)
plot.loess<-vector("list", max(msd01$Rank_T21byAge))
plot.linear<-vector("list", max(msd01$Rank_T21byAge))
for (i in 1:max(msd01$Rank_T21byAge)) {
  df<-msd01 %>% filter(Rank_T21byAge==i)
  analyte<-df$Analyte[1]
  plot.loess_i<-ggplot(aes(x=Age, y=Concentration, colour=Cohort), data=df) +
    geom_smooth() +
#    geom_smooth(data=sim.slopeAge_SAA, aes(x=Age, y = Concentration), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
    ggtitle(analyte) +
    scale_fill_manual(values=RedBlue) +
    scale_color_manual(values=RedBlue)
  plot.linear_i<-ggplot(aes(x=Age, y=Concentration, colour=Cohort), data=df) +
    geom_smooth(method="lm") +
    ggtitle(analyte) +
    scale_fill_manual(values=RedBlue) +
    scale_color_manual(values=RedBlue) #+

    setwd(paste(out.dir, "PNG"))
    ggsave(filename = paste0(Analyte_filename, ".png", sep=""), width = 3.5, height = 4, units = "in")
    setwd(paste(out.dir, "PDF"))
    ggsave(filename = paste0(Analyte_filename, ".pdf", sep=""), device = cairo_pdf, width = 3.5, height = 4, units = "in")
  
  plot.linear[[i]]<-plot.linear_i
  plot.loess[[i]]<-plot.loess_i
}

length(plot.loess)

# setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Age")
# pdf("200503_Concentration_by_Age_LOESS.pdf")
# plot.loess
# dev.off()
# 
# setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Age")
# pdf("200503_Concentration_by_Age_Linear.pdf")
# print(plot.linear)
# dev.off()
```

# Now with log2 concentration
```{r}
plot.loess<-vector("list", 54)
plot.linear<-vector("list", 54)
index<-msd %>% select(Analyte) %>% unique() %>% mutate(Index=seq(1:54))

msd01<-msd %>% full_join(index, by="Analyte") %>%
  mutate(Cohort=ifelse(Cohort=="Control", "D21",
                       ifelse(Cohort=="Down syndrome", "T21", NA)))

for (i in 1:54) {
  df<-msd01 %>% filter(Index==i)
  analyte<-df$Analyte[1]
  plot.loess[[i]]<-ggplot(aes(x=Age, y=log2(Concentration), colour=Cohort), data=df) +
    geom_smooth() +
    ggtitle(analyte) +
    scale_fill_manual(values=RedBlue) +
    scale_color_manual(values=RedBlue)
  plot.linear[[i]]<-ggplot(aes(x=Age, y=log2(Concentration), colour=Cohort), data=df) +
    geom_smooth(method="lm") +
    ggtitle(analyte) +
    scale_fill_manual(values=RedBlue) +
    scale_color_manual(values=RedBlue) #+
    #theme(aspect.ratio = 0.8,
    #    legend.position = c(0.1, 0.85)) # depends on data scales
        #legend.position = "none" ) +
   #    legend.background = element_blank() )
}

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Age")
pdf("200503_log2Concentration_by_Age_LOESS.pdf")
plot.loess
dev.off()

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Age")
pdf("200503_log2Concentration_by_Age_Linear.pdf")
print(plot.linear)
dev.off()
```









#### Plot predicted concentration over age for cytokines with significant T21*Age interactions
##### SAA
```{r}
# Setting and modifying theme for plots
theme_set(theme_gray(base_size = 12, base_family = "Arial") +
            theme(panel.border = element_rect(colour="black", fill = "transparent"),
                  plot.title = element_text(face="bold", hjust = 0), # lineheight=.8, size=20,
                  axis.text = element_text(color="black", size = 14), 
                  axis.text.x = element_text(angle = 0, hjust = NULL),
                  strip.background = element_rect(colour="black", fill = "light grey", size = 1), # adjusts facet label borders (if any)
                  panel.background = element_blank(),
                  panel.grid = element_blank()
          ))

# lm
ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="VascularInjury_SAA"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="lm") +
  geom_smooth(aes(x=Age, y = .fixed), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.1, 0.85)) + # depends on data scales
        #legend.position = "none" ) +
   #    legend.background = element_blank() ) +
  labs(title = "Predicted Serum Amyloid A (SAA) over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+(1|Source)")
setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
#ggsave(filename = "200430_MSD_PredictedOverAge_Linear_SAA_Model2.png", width = 5, height = 5, units = "in")
#ggsave("200430_MSD_PredictedOverAge_Linear_SAA_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")

# loess
ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="VascularInjury_SAA"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="loess") +
  geom_smooth(aes(x=Age, y = .fixed), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.1, 0.85)) + # depends on data scales
        #legend.position = "none" ) +
   #    legend.background = element_blank() ) +
  labs(title = "Predicted Serum Amyloid A (SAA) over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+(1|Source)")
setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
#ggsave(filename = "200430_MSD_PredictedOverAge_LOESS_SAA_Model2.png", width = 5, height = 5, units = "in")
#ggsave("200430_MSD_PredictedOverAge_LOESS_SAA_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")
```

##### UPLEX_IL-29	
```{r}
# Setting and modifying theme for plots
theme_set(theme_gray(base_size = 12, base_family = "Arial") +
            theme(panel.border = element_rect(colour="black", fill = "transparent"),
                  plot.title = element_text(face="bold", hjust = 0), # lineheight=.8, size=20,
                  axis.text = element_text(color="black", size = 14), 
                  axis.text.x = element_text(angle = 0, hjust = NULL),
                  strip.background = element_rect(colour="black", fill = "light grey", size = 1), # adjusts facet label borders (if any)
                  panel.background = element_blank(),
                  panel.grid = element_blank()
          ))

ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="UPLEX_IL-29"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="lm") +
  geom_smooth(aes(x=Age, y = .fitted), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
#  ylab("Predicted log(Concentration)") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
# ylab(paste("Predicted ", bquote(''~log[2](Concentration)~''), sep="" )) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  #ylim(1,4) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.15, 0.95), # depends on data scales
        legend.background = element_blank(),
        legend.title=element_blank(),
        legend.box = "horizontal") +
  labs(title = "Predicted IL-29 over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+T21*Age+(1|Source)") +
  guides(colour = guide_legend(nrow = 1))


setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
ggsave(filename = "200430_MSD_PredictedOverAge_Linear_IL29_Model2.png", width = 5, height = 5, units = "in")
ggsave("200430_MSD_PredictedOverAge_Linear_IL29_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")

ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="UPLEX_IL-29"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="loess") +
  geom_smooth(aes(x=Age, y = .fitted), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
#  ylab("Predicted log(Concentration)") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
# ylab(paste("Predicted ", bquote(''~log[2](Concentration)~''), sep="" )) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  #ylim(1,4) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.15, 0.95), # depends on data scales
        legend.background = element_blank(),
        legend.title=element_blank(),
        legend.box = "horizontal") +
  labs(title = "Predicted IL-29 over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+T21*Age+(1|Source)") +
  guides(colour = guide_legend(nrow = 1))

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
ggsave(filename = "200430_MSD_PredictedOverAge_LOESS_IL29_Model2.png", width = 5, height = 5, units = "in")
ggsave("200430_MSD_PredictedOverAge_LOESS_IL29_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")
```

##### TH17_IL-22

##### TH17_IL-27


##### Angiogenesis_VEGF-C
```{r}
# Setting and modifying theme for plots
theme_set(theme_gray(base_size = 12, base_family = "Arial") +
            theme(panel.border = element_rect(colour="black", fill = "transparent"),
                  plot.title = element_text(face="bold", hjust = 0), # lineheight=.8, size=20,
                  axis.text = element_text(color="black", size = 14), 
                  axis.text.x = element_text(angle = 0, hjust = NULL),
                  strip.background = element_rect(colour="black", fill = "light grey", size = 1), # adjusts facet label borders (if any)
                  panel.background = element_blank(),
                  panel.grid = element_blank()
          ))

ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="VascularInjury_SAA"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="lm") +
  geom_smooth(aes(x=Age, y = .fitted), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
#  ylab("Predicted log(Concentration)") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
# ylab(paste("Predicted ", bquote(''~log[2](Concentration)~''), sep="" )) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.1, 0.85)) + # depends on data scales
        #legend.position = "none" ) +
   #    legend.background = element_blank() ) +
  labs(title = "Predicted Serum Amyloid A (SAA) over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+T21*Age+(1|Source)")

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
ggsave(filename = "200430_MSD_PredictedOverAge_Linear_SAA_Model2.png", width = 5, height = 5, units = "in")
ggsave("200430_MSD_PredictedOverAge_Linear_SAA_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")


ggplot(data=augment %>%
         mutate(Cohort=ifelse(T21==1, "T21", ifelse(T21==0, "D21", NA))) %>%
         filter(Panel_Analyte=="VascularInjury_SAA"),
       aes(x=Age, y = .fitted, color=Cohort)) +
  geom_smooth(method="loess") +
  geom_smooth(aes(x=Age, y = .fitted), color=Purple, method="lm", se = FALSE, linetype = "dashed") +
#  ylab("Predicted log(Concentration)") +
  ylab(bquote('Predicted'~log[2](Concentration)~'')) +
# ylab(paste("Predicted ", bquote(''~log[2](Concentration)~''), sep="" )) +
  scale_fill_manual(values=RedBlue) +
  scale_color_manual(values=RedBlue) +
  theme(aspect.ratio = 0.8,
        legend.position = c(0.1, 0.85)) + # depends on data scales
        #legend.position = "none" ) +
   #    legend.background = element_blank() ) +
  labs(title = "Predicted Serum Amyloid A (SAA) over Age",
    subtitle = "LMM: log2(Conc.)~T21+Age+Female+T21*Age+(1|Source)")

setwd("/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/MSD/Final/200430/Plots/Predicted_over_age")
ggsave(filename = "200430_MSD_PredictedOverAge_LOESS_SAA_Model2.png", width = 5, height = 5, units = "in")
ggsave("200430_MSD_PredictedOverAge_LOESS_SAA_Model2.pdf", device = cairo_pdf, width = 5, height = 5, units = "in")
```

