---
title: "R Notebook"
output: html_notebook
---

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata for clustering among D21s of protein trajectories with significant T21*Age interaction
```{r}
setwd(dir.ardata)
ar_p4c_soma_sigt21byage_d21only_clustering_input <- fread("ar_p4c_soma_sigt21byage_d21only_clustering_input.csv.gz")

head(ar_p4c_soma_sigt21byage_d21only_clustering_input)
```

#### Split ardata by UniProt_Aptamer_Chr_Cluster (used as uid used for convenience)
```{r}
ar_p4c_soma_sigt21byage_d21only_clustering_input.split <- ar_p4c_soma_sigt21byage_d21only_clustering_input %>%
  split(., .$UniProt_Aptamer_Chr_Cluster);

ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[1]] 
```


```{r}
# https://purrr.tidyverse.org/reference/quietly.html
quiet_loess <- purrr::quietly(loess)
```

#### For each aptamer with significant T21*Age interaction, obtain Loess predicted values for each D21
```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.loess_prediction.D21.df <- list()
loess_prediction.D21 <- list()
#plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_sigt21byage_d21only_clustering_input.split) ){
#for ( i in 1:3 ){
  set.seed(1234)
  tmp.loess_prediction.D21.df[[i]] <- ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age + Female,
                                                        # IMPORTANT: ADJUST FOR SEX FOR BETTER CLLUSTERING?
                                                        data = ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]]))$fitted);
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  #names(loess_prediction.D21)[[i]] <- names(ar_p4c_soma_d21only_extend_clustering.split)[[i]]
  names(tmp.loess_prediction.D21.df)[[i]] <- names(ar_p4c_soma_sigt21byage_d21only_clustering_input.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()
#tmp.loess_prediction.D21.df[[1]] %>% head()

loess_prediction.D21.df <- tmp.loess_prediction.D21.df %>%
  rbindlist();

head(loess_prediction.D21.df)
#rm(tmp.loess_prediction.D21.df); gc();

loess_prediction.D21.matrix <- tmp.loess_prediction.D21.df %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, predicted_Zlog10_relative_abundance) %>%
  unique() %>%
  spread(key = LabID, value = predicted_Zlog10_relative_abundance) %>%
  column_to_rownames(var = "UniProt_Aptamer_Chr_Cluster") %>%
  as.matrix();
#predicted.D21.matrix %>% as.data.frame()

# Dissimilarity matrix:
set.seed(1234)
d_newclusters.D21s <- dist(loess_prediction.D21.matrix,
                           method = "euclidean");
# Hierarchical clustering using Complete Linkage:
set.seed(1234)
hc1_newclusters.D21s <- hclust(d_newclusters.D21s,
                               method = "complete");
```

#### Decide how many T21-by-Age Clusters make sense
```{r}
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 4, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 8, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 9, border = 1:9);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 10, border = 1:10);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 11, border = 1:11);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 12, border = 1:12);

k_hc1_newclusters.D21s <- 12
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s, border = 1:k_hc1_newclusters.D21s);
```

#### Use the cuttree function to assign aptamers to k=12 T21-by-Age Clusters ('ar_sigt21byage_d21clustered')
```{r}
k_hc1_newclusters.D21s <- 12

ar_sigt21byage_d21clustered <- cutree(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s) %>%
    as.data.frame() %>%
    `colnames<-`("T21-by-Age Cluster") %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
    mutate(k = k_hc1_newclusters.D21s) %>%
    ungroup() %>%
  left_join( fread(paste0(dir.ardata, "/ar_p4c_soma.csv.gz")), by = "UniProt_Aptamer_Chr_Cluster" )

head(ar_sigt21byage_d21clustered)

# ar_sigt21byage_d21clustered
```

```{r}
setwd(dir.ardata)
fwrite(ar_sigt21byage_d21clustered, "ar_sigt21byage_d21clustered.csv.gz")
```

