---
title: "R Notebook"
output: html_notebook
---

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata
```{r}
setwd(dir.ardata)
ar_p4c_soma_d21only <- fread("ar_p4c_soma_d21only.csv.gz") #%>%
  #filter(Aptamer_in_P4C == 1 & Aptamer_in_Lehallier2019 == 0);

#ar_p4c_soma_d21only
#ar_p4c_soma_t21only
```

#### Review the overlap of SomaLogic Aptamers analyzed in Lehallier et al., 2019 vs. in the P4C
```{r}
ar_p4c_soma_d21only %>%
  dplyr::select(Aptamer, Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>%
  unique() %>%
  group_by(Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>% 
  summarise(N = n())

# ar_p4c_soma_t21only %>%
#   dplyr::select(Aptamer, Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>%
#   unique() %>%
#   group_by(Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>% 
#   summarise(N = n())
```

```{r}
ar_p4c_soma_d21only.split <- ar_p4c_soma_d21only %>%
  # WARNING: TESTING THIS ONLY, REMEMBER TO REMOVE NEXT 2 LINES LATER IF NOT THE BEST APPROACH:
  filter(Age>=18) %>%
  filter(Aptamer_in_Lehallier2019 == 1) %>%
  split(., .$UniProt_Aptamer_Chr_Cluster);

rm(ar_p4c_soma_d21only); gc()
```

```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
predicted_values.D21 <- list()
plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_d21only.split) ){
  set.seed(1234)
  loess_newclusters.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
                          data = ar_p4c_soma_d21only.split[[i]])
  
  predicted_values.D21[[i]] <- cbind(ar_p4c_soma_d21only.split[[i]]$LabID,
                                        loess_newclusters.D21[[i]]$fitted,
                                        ar_p4c_soma_d21only.split[[i]]$Zlog10_relative_abundance,
                                        ar_p4c_soma_d21only.split[[i]]$Age,
                                        as.character(ar_p4c_soma_d21only.split[[i]]$Sex)) %>%
    as.data.frame() %>%
    `colnames<-`(c("LabID", "predicted_Zlog10_relative_abundance", "Zlog10_relative_abundance", "Age", "Sex")) %>%
    mutate(Age = as.numeric(Age),
           predicted_Zlog10_relative_abundance = as.numeric(predicted_Zlog10_relative_abundance),
           Zlog10_relative_abundance = as.numeric(Zlog10_relative_abundance),
           Sex = factor(Sex, levels = c("Male", "Female"))) %>%
    arrange(Age) %>%
    mutate(UniProt_Aptamer_Chr_Cluster = names(ar_p4c_soma_d21only.split)[[i]]) %>%
    dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, Age, Sex,
                  predicted_Zlog10_relative_abundance, Zlog10_relative_abundance) %>%
    separate(UniProt_Aptamer_Chr_Cluster, into = c("UniProt", "Other"), sep = "[|]", extra = "merge", remove = FALSE) %>%
    separate(Other, into = c("Aptamer", "Other"), sep = " [(]", extra = "merge", remove = TRUE) %>%
    separate(Other, into = c("Chromosome", "Lehallier2019_ClusterID"), sep = "[,] ", extra = "merge", remove = TRUE) %>%
    mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID),
           Lehallier2019_ClusterID = trimws(Lehallier2019_ClusterID),
           Chromosome = trimws(Chromosome),
           UniProt = trimws(UniProt),
           Aptamer = trimws(Aptamer)) %>%
  separate(UniProt_Aptamer_Chr_Cluster, into = c("UniProt", "Other"), sep = "[|]", extra = "merge", remove = FALSE) %>%
  separate(Other, into = c("Aptamer", "Other"), sep = " [(]", extra = "merge", remove = TRUE) %>%
  separate(Other, into = c("Chromosome", "Lehallier2019_ClusterID"), sep = "[,] ", extra = "merge", remove = TRUE) %>%
  mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID),
         Lehallier2019_ClusterID = trimws(Lehallier2019_ClusterID),
         Chromosome = trimws(Chromosome),
         UniProt = trimws(UniProt),
         Aptamer = trimws(Aptamer)) %>%
  group_by(LabID, Age, Sex, Lehallier2019_ClusterID) %>%
  summarise(median_predicted_Zlog10_relative_abundance = median(predicted_Zlog10_relative_abundance)) %>%
  ungroup() %>%
  mutate(UniProt_Aptamer_Chr_Cluster = paste0(Lehallier2019_ClusterID, " representative"),
         type_Zlog10_relative_abundance = "Median predicted value for LabID in Cluster") %>%
  dplyr::rename(Zlog10_relative_abundance = median_predicted_Zlog10_relative_abundance) %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, LabID, Age, Sex,
         Zlog10_relative_abundance, type_Zlog10_relative_abundance)
}
#rm(loess_newclusters.D21, ar_p4c_soma_d21only.split); gc()

tmp1 <- predicted_values.D21 %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, LabID, Age, Sex,
         Zlog10_relative_abundance, type_Zlog10_relative_abundance)
setwd(dir.ardata)
tmp2 <- fread("ar_p4c_soma_d21only.csv.gz") %>%
  filter(Age>=18) %>% # IMPORTANT: MUST FILTER HERE AS ABOVE BEFORE INPUT TO PREDICTION CODE.
  #filter(Aptamer_in_Lehallier2019 == 0) %>% # ie, keep aptamers not already clustered by Lehallier 2019
  mutate(type_Zlog10_relative_abundance = "Actual value") %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, LabID, Age, Sex,
         Zlog10_relative_abundance, type_Zlog10_relative_abundance)

ar_p4c_soma_d21only_extend_clustering <- rbind(tmp1, tmp2);

ar_p4c_soma_d21only_extend_clustering

setwd(dir.ardata)
fwrite(ar_p4c_soma_d21only_extend_clustering, "ar_p4c_soma_d21only_extend_clustering.csv.gz")
```

#### Within D21s only, cluster the n=2607 proteins already clustered by Lehallier 2019 to obtain loess predictions and derive the median loess predicted values for each LabID within each Cluster
```{R}
ar_p4c_soma_d21only_extend_clustering.split <- ar_p4c_soma_d21only_extend_clustering %>%
  split(., .$UniProt_Aptamer_Chr_Cluster)

length(ar_p4c_soma_d21only_extend_clustering.split) #[1] 4640

# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.loess_prediction.D21.df <- list()
loess_prediction.D21 <- list()
#plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_d21only_extend_clustering.split) ){
#for ( i in 1:3 ){
  set.seed(1234)

  #loess_prediction.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
  #                                    data = ar_p4c_soma_d21only_extend_clustering.split[[i]])

  tmp.loess_prediction.D21.df[[i]] <- ar_p4c_soma_d21only_extend_clustering.split[[i]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age,
                                                        data = ar_p4c_soma_d21only_extend_clustering.split[[i]]))$fitted)
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  #names(loess_prediction.D21)[[i]] <- names(ar_p4c_soma_d21only_extend_clustering.split)[[i]]
  names(tmp.loess_prediction.D21.df)[[i]] <- names(ar_p4c_soma_d21only_extend_clustering.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()

loess_prediction.D21.df <- tmp.loess_prediction.D21.df %>%
  rbindlist();
#rm(tmp.loess_prediction.D21.df); gc();

loess_prediction.D21 %>% lapply(class) %>% unique()

loess_prediction.D21.matrix <- tmp.loess_prediction.D21.df %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, predicted_Zlog10_relative_abundance) %>%
  unique() %>%
  spread(key = LabID, value = predicted_Zlog10_relative_abundance) %>%
  column_to_rownames(var = "UniProt_Aptamer_Chr_Cluster") %>%
  as.matrix();
#predicted.D21.matrix %>% as.data.frame()

# Dissimilarity matrix:
set.seed(1234)
d_newclusters.D21s <- dist(loess_prediction.D21.matrix,
                           method = "euclidean");
# Hierarchical clustering using Complete Linkage:
set.seed(1234)
hc1_newclusters.D21s <- hclust(d_newclusters.D21s,
                               method = "complete");
```

```{r}
# Cut tree into 6 groups
k.now <- 8; # Best so far = 28
sub_grp6_newclusters.D21s <- cutree(hc1_newclusters.D21s, k = k.now)
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k.now, border = 1:k.now);

# Add the the cluster each observation belongs to to our original data.
# length(sub_grp6_newclusters.D21s) #[1] 4632
# predicted_newclusters.D21.matrix %>% as.data.frame() %>% nrow() #[1] 4632
```

#### ai_d21_extended_clustering
```{r}
#rm(d21_cluster_mapping, d21_clustering); gc()
tmp.d21_extended_clustering <- list()
d21_cluster_mapping <- list()
for (i in seq(1:40)) {
  # Note: Best k so far = 28
  
  #k.now <- i + 7
  k.now <- i
  
  print(k.now)
  set.seed(1234)

  # Cut tree into k groups
  tmp.d21_extended_clustering[[i]] <- cutree(hc1_newclusters.D21s, k = k.now) %>%
    as.data.frame() %>%
    `colnames<-`("P4C_Clock_Cluster") %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
    mutate(k = k.now) %>%
    ungroup();
  #plot(hc1_newclusters.D21s, cex = 0.1);
  #rect.hclust(hc1_newclusters.D21s, k = k.now, border = 1:k.now);
}

tmp.d21_extended_clustering %>% length() # 23
#tmp.d21_extended_clustering %>% head(n=3)

#ar_p4c_soma_d21only_extend_clustering.split %>% rbindlist()
#tmp.d21_extended_clustering %>% rbindlist()

ai_d21_extended_clustering <- tmp.d21_extended_clustering %>%
  rbindlist() %>%
  mutate(tmp = gsub("Cluster ", "~", UniProt_Aptamer_Chr_Cluster)) %>%
  separate(tmp, into = c("Other", "Lehallier2019_ClusterID"),
            sep = "[~]",
            extra = "merge",
            remove = TRUE) %>%
  mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID)) %>%
  select(k, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  unique() %>%
  `rownames<-`(NULL) %>%
  mutate(Lehallier2019_ClusterID = gsub(" representative", "", Lehallier2019_ClusterID),
         Lehallier2019_ClusterID = gsub("Other", "None", Lehallier2019_ClusterID)) %>%
  select(k, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster);

setwd(dir.aidata)
fwrite(ai_d21_extended_clustering, "ai_d21_extended_clustering.csv.gz")

head(ai_d21_extended_clustering)
```

```{r}
ai_d21_extended_clustering

ai_d21_extended_cluster_mapping <- ai_d21_extended_clustering %>%
  filter(grepl("epresentative", UniProt_Aptamer_Chr_Cluster)==TRUE)
```

```{r}
ai_d21_extended_cluster_mapping <- ai_d21_extended_clustering %>%
  filter(grepl("epresentative", UniProt_Aptamer_Chr_Cluster)==TRUE) %>%
  #filter(k == 8)  %>%
  arrange(Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  dplyr::select(k, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  unique() %>%
  arrange(Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  group_by(k, UniProt_Aptamer_Chr_Cluster) %>%
  mutate(P4C_Clock_Cluster = paste0("P4C_Clock_Cluster_", P4C_Clock_Cluster),
         Lehallier2019_ClusterID = paste0("Lehallier2019_Cluster_", Lehallier2019_ClusterID))

ai_d21_extended_cluster_mapping %>%
  filter(k>=8 & k<=20) %>%
  select(-c(UniProt_Aptamer_Chr_Cluster)) %>%
  split(., .$k)
```

```{r}
ai_d21_extended_cluster_mapping %>%
  spread(key = Lehallier2019_ClusterID, value = P4C_Clock_Cluster) %>%
  ungroup() %>%
  dplyr::select(-c(UniProt_Aptamer_Chr_Cluster)) %>%
  gather(key = "key", value = "value", starts_with("Lehallier")) %>%
  filter(!is.na(value)) %>%
  group_by(k, key) %>%
  summarise(N_Lehallier2019_Clusters = n()) %>%
  ungroup() %>%
  filter(k>=8) %>%
  group_by(k, key) %>%
  summarise(max.N_Lehallier2019_Clusters = max(N_Lehallier2019_Clusters))
```

```{r}
ai_d21_extended_cluster_mapping %>%
  gather(key = "key", value = "value", starts_with("Lehallier")) %>%
  filter(!is.na(value)) %>%
  arrange(k, key, value) %>%
  group_by(k, key) %>%
  summarise(N_Lehallier2019_Clusters = n()) %>%
  ungroup() %>%
  group_by(k) %>%
  summarise(max.N_Lehallier2019_Clusters = max(N_Lehallier2019_Clusters))

ai_d21_extended_cluster_mapping %>%
  gather(key = "key", value = "value", starts_with("P4C_Clock_Cluster_")) %>%
  filter(!is.na(value)) %>%
  group_by(k, key) %>%
  

```


```{r}
ai_d21_extended_clustering %>%
  filter(grepl("epresentative", UniProt_Aptamer_Chr_Cluster)==TRUE) %>%
  #filter(k == 8)  %>%
  arrange(Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  select(k, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  unique() %>%
  group_by(k, P4C_Clock_Cluster) %>%
  summarise(N_P4C_Clock_Cluster_mappings = n()) %>%
  ungroup() %>%
  group_by(k, N_P4C_Clock_Cluster_mappings) %>%
  summarise(N = n()) %>%
  ungroup() %>%
  group_by(k) %>%
  spread(key = N_P4C_Clock_Cluster_mappings, value = N)


  summarise(max_N_P4C_Clock_Cluster_mappings = max(N_P4C_Clock_Cluster_mappings)) %>%
  ungroup() %>%
  arrange(max_N_P4C_Clock_Cluster_mappings)
```

```{r}
  d21_cluster_mapping[[i]] <- predicted_newclusters.D21.matrix %>%
    as.data.frame() %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
    full_join(d21_clustering[[i]], by = "UniProt_Aptamer_Chr_Cluster") %>%
    #mutate(P4C_Clock_Cluster = d21_clustering[[i]]) %>%
    dplyr::select(P4C_Clock_Cluster, everything()) %>%
    gather(key = "LabID", value = "predicted_Zlog10_relative_abundance", starts_with("HTP")) %>%
    ungroup() %>%
    filter(grepl("Lehallier", UniProt_Aptamer_Chr_Cluster)==TRUE) %>%
    mutate(Lehallier2019_ClusterID = case_when(grepl("Lehallier", UniProt_Aptamer_Chr_Cluster)==TRUE ~
                                                 as.numeric(gsub("_representative", "",
                                                      gsub("Lehallier2019_Cluster", "", UniProt_Aptamer_Chr_Cluster))),
                                               .default = NA),
           P4C_Clock_Cluster = as.numeric(P4C_Clock_Cluster)) %>%
    dplyr::select(k, Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
    unique() %>%
    `rownames<-`(NULL) %>%
    arrange(P4C_Clock_Cluster)
}

length(d21_cluster_mapping)

d21_cluster_mapping %>%
  rbindlist() %>%
  split(., .$Lehallier2019_ClusterID)
```




```{r}
setwd(dir.ardata)
ar_p4c_soma_d21only_P4ConlyAptamers <- fread("ar_p4c_soma_d21only.csv.gz") %>%
  filter(Aptamer_in_P4C==1 & Aptamer_in_Lehallier2019==0);

set.seed(1234)
predicted_newclusters.D21.df <- predicted_newclusters.D21 %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, Age, Sex,
                predicted_Zlog10_relative_abundance, Zlog10_relative_abundance) %>%
  unique() %>%
  separate(UniProt_Aptamer_Chr_Cluster, into = c("UniProt", "Other"), sep = "[|]", extra = "merge", remove = FALSE) %>%
  separate(Other, into = c("Aptamer", "Other"), sep = " [(]", extra = "merge", remove = TRUE) %>%
  separate(Other, into = c("Chromosome", "Lehallier2019_ClusterID"), sep = "[,] ", extra = "merge", remove = TRUE) %>%
  mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID),
         Lehallier2019_ClusterID = trimws(Lehallier2019_ClusterID),
         Chromosome = trimws(Chromosome),
         UniProt = trimws(UniProt),
         Aptamer = trimws(Aptamer)) %>%
  dplyr::select(-c(Chromosome, UniProt, Aptamer)) %>%
  unique() %>%
  group_by(LabID, Age, Sex, Lehallier2019_ClusterID) %>%
  summarise(N_obs = n(),
            median_predicted_Zlog10_relative_abundance = median(predicted_Zlog10_relative_abundance),
            median_Zlog10_relative_abundance = median(Zlog10_relative_abundance)) %>%
  ungroup() %>%
  mutate(UniProt_Aptamer_Chr_Cluster = paste0("Lehallier2019_",
                          gsub(" ", "", Lehallier2019_ClusterID), "_representative"));
  
#rm(predicted_newclusters.D21); gc()

head(predicted_newclusters.D21.df)

predicted_newclusters.D21.df %>%
    mutate(type_Zlog10_relative_abundance = "Median age-predicted value for LabID within Cluster") %>%
    dplyr::rename(Zlog10_relative_abundance = median_predicted_Zlog10_relative_abundance) %>%
    dplyr::select(LabID, Age, Sex,
                  UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID,
                  Zlog10_relative_abundance, type_Zlog10_relative_abundance)
  
  
ar_p4c_soma_d21only_extend_clustering <- rbind( ),
                                                (ar_p4c_soma_d21only_P4ConlyAptamers %>%
                                                  mutate(type_Zlog10_relative_abundance = "Actual value for LabID") %>%
                                                  dplyr::select(LabID, Age, Sex, UniProt_Aptamer_Chr_Cluster,
                                                                Lehallier2019_ClusterID,
                                                                Zlog10_relative_abundance, type_Zlog10_relative_abundance)) )

ar_p4c_soma_d21only_extend_clustering
```


```{r}
predicted_newclusters.D21.matrix %>%
    as.data.frame() %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
  gather(key = "LabID", value = "predicted_Zlog10_relative_abundance", starts_with("HTP"))
```



```{r}
tmp <- predicted_newclusters.D21.matrix %>%
  as.data.frame() %>%
  #mutate(P4C_Clock_Cluster = 8 + sub_grp6_newclusters.D21s) %>% # Assign cluster numbers 9-16 since Lehallier 2019 defined clusters 1-8. # First option tried...
    mutate(P4C_Clock_Cluster = sub_grp6_newclusters.D21s) %>%
  dplyr::select(P4C_Clock_Cluster, everything()) %>%
  rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
  gather(key = "LabID", value = "predicted_Zlog10_relative_abundance", starts_with("HTP")) %>%
  ungroup() %>%
  left_join( (ar_p4c_soma_d21only %>%
                 dplyr::select(LabID, Age, Sex, Aptamer, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID) %>%
                 # WARNING: NEXT LINE IS FOR TESTING - REMOVE IF NOT BEST APPROACH:
                 filter(Age>=18 ^ !is.na(Age)) %>%
                 unique() %>%
                 ungroup()),
              by = c("LabID", "UniProt_Aptamer_Chr_Cluster") ) %>%
  dplyr::select(LabID, Age, Sex, Aptamer, UniProt_Aptamer_Chr_Cluster, 
         Lehallier2019_ClusterID, P4C_Clock_Cluster, contains("Cluster"),
         predicted_Zlog10_relative_abundance, everything()) %>%
#  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID,
#                P4C_Clock_Cluster, everything()) %>%
  unique() %>%
  ungroup();

tmp %>%
  group_by(Lehallier2019_ClusterID, P4C_Clock_Cluster)

```


```{r}
ai_p4c_soma_d21only_aging_clusters <- predicted_newclusters.D21.matrix %>%
  as.data.frame() %>%
  #mutate(P4C_Clock_Cluster = 8 + sub_grp6_newclusters.D21s) %>% # Assign cluster numbers 9-16 since Lehallier 2019 defined clusters 1-8. # First option tried...
    mutate(P4C_Clock_Cluster = sub_grp6_newclusters.D21s) %>% # New option from clustering ALL measured proteins.
  dplyr::select(P4C_Clock_Cluster, everything()) %>%
  rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
  gather(key = "LabID", value = "predicted_Zlog10_relative_abundance", starts_with("HTP")) %>%
  ungroup() %>%
  mutate(LabID = trimws(LabID),
         UniProt_Aptamer_Chr_Cluster = trimws(UniProt_Aptamer_Chr_Cluster)) %>%
  left_join( (ar_p4c_soma_d21only %>%
                 dplyr::select(LabID, Age, Sex, Aptamer, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID) %>%
                 # WARNING: NEXT LINE IS FOR TESTING - REMOVE IF NOT BEST APPROACH:
                 filter(Age>=18) %>%
                 unique() %>%
                 filter(!is.na(Age)) %>%
                 ungroup() %>%
                 mutate(LabID = trimws(LabID),
                        UniProt_Aptamer_Chr_Cluster = trimws(UniProt_Aptamer_Chr_Cluster))),
              by = c("LabID", "UniProt_Aptamer_Chr_Cluster") ) %>%
  dplyr::select(LabID, Age, Sex, Aptamer, UniProt_Aptamer_Chr_Cluster, 
         Lehallier2019_ClusterID, P4C_Clock_Cluster, contains("Cluster"),
         predicted_Zlog10_relative_abundance, everything()) %>%
#  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID,
#                P4C_Clock_Cluster, everything()) %>%
  unique() %>%
  ungroup();

#setwd(dir.ardata)
#fwrite(ai_p4c_soma_d21only_aging_clusters, "ai_p4c_soma_d21only_aging_clusters.csv.gz")
```

```{r}
tmp <- ai_p4c_soma_d21only_aging_clusters %>%
  dplyr::select(Aptamer, Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  unique() %>%
  group_by(Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  summarise(N_Aptamers = n()) %>%
  ungroup() %>%
  arrange(Lehallier2019_ClusterID, P4C_Clock_Cluster) %>%
  group_by(Lehallier2019_ClusterID) %>%
  mutate(N_Aptamers_Total = sum(N_Aptamers),
         Proportion_of_Lehallier2019_Cluster = N_Aptamers/N_Aptamers_Total) %>%
  ungroup() %>%
  mutate(#Lehallier2019_ClusterID = paste0("Lehallier_", Lehallier2019_ClusterID),
         Lehallier2019_ClusterID = factor(Lehallier2019_ClusterID, levels = c(as.character(seq(1:8)))),
         P4C_Clock_Cluster = paste0("P4C_Cluster", P4C_Clock_Cluster),
         P4C_Clock_Cluster = factor(P4C_Clock_Cluster,
                                    levels = c(paste0("P4C_Cluster", seq(1:8)))));

tmp %>%
  group_by(Lehallier2019_ClusterID) %>%
  arrange(desc(Proportion_of_Lehallier2019_Cluster)) %>%
  mutate(Proportion_rank = row_number()) %>%
  arrange(Proportion_rank) %>%
  ungroup()

gg_p4c_soma_aging_cluster_barplot <- tmp %>%
  dplyr::rename(Lehallier2019_Cluster = Lehallier2019_ClusterID) %>%
  ggplot(aes(Lehallier2019_Cluster, y = Proportion_of_Lehallier2019_Cluster,
             color = P4C_Clock_Cluster, fill = P4C_Clock_Cluster)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylab("Proportion\nof Lehallier Cluster") +
  theme(aspect.ratio = 0.5,
        legend.position = "bottom",
        legend.title = element_blank()) #+
  #facet_wrap(~Lehallier2019_ClusterID, ncol=4);

gg_p4c_soma_aging_cluster_barplot
```
