---
title: "R Notebook"
output: html_notebook
---

# KEEP SCRIPT

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata for clustering among D21s of protein trajectories with significant T21*Age interaction
```{r}
setwd(dir.ardata)
ar_p4c_soma_sigt21byage_d21only_clustering_input <- fread("ar_p4c_soma_sigt21byage_d21only_clustering_input.csv.gz")

setwd(dir.ardata)
ar_p4c_soma_sigt21byage_t21only_clustering_input <- fread("ar_p4c_soma_sigt21byage_t21only_clustering_input.csv.gz")

head(ar_p4c_soma_sigt21byage_d21only_clustering_input)
head(ar_p4c_soma_sigt21byage_t21only_clustering_input)
```

#### Split ardata by UniProt_Aptamer_Chr_Cluster (used as uid used for convenience)
```{r}
ar_p4c_soma_sigt21byage_d21only_clustering_input.split <- ar_p4c_soma_sigt21byage_d21only_clustering_input %>%
  split(., .$UniProt_Aptamer_Chr_Cluster);

ar_p4c_soma_sigt21byage_t21only_clustering_input.split <- ar_p4c_soma_sigt21byage_t21only_clustering_input %>%
  split(., .$UniProt_Aptamer_Chr_Cluster);

ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[1]]
ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[1]]

ar_p4c_soma_sigt21byage_d21only_clustering_input$UniProt_Aptamer_Chr_Cluster %>% unique() %>% length()
ar_p4c_soma_sigt21byage_t21only_clustering_input$UniProt_Aptamer_Chr_Cluster %>% unique() %>% length()
#[1] 850
#[1] 850
```

```{r}
# https://purrr.tidyverse.org/reference/quietly.html
quiet_loess <- purrr::quietly(loess)
```

```{r}
# loess(formula, data, weights, subset, na.action, model = FALSE,
#       span = 0.75, enp.target, degree = 2,
#       parametric = FALSE, drop.square = FALSE, normalize = TRUE,
#       family = c("gaussian", "symmetric"),
#       method = c("loess", "model.frame"),
#       control = loess.control(...), ...)
# 
# parametric - should any terms be fitted globally rather than locally? Terms can be specified by name, number or as a logical vector of the same length as the number of predictors.
# 
# model	- should the model frame be returned?
#   
# degree - the degree of the polynomials to be used, normally 1 or 2. (Degree 0 is also allowed, but see the ‘Note’.)
# 
# family - if "gaussian" fitting is by least-squares, and if "symmetric" a re-descending M estimator is used with Tukey's biweight function. Can be abbreviated.
# 
# method - fit the model or just extract the model frame. Can be abbreviated.
# 
# control - control parameters: see loess.control.


# loess.control(surface = c("interpolate", "direct"),
#               statistics = c("approximate", "exact", "none"),
#               trace.hat = c("exact", "approximate"),
#               cell = 0.2, iterations = 4, iterTrace = FALSE, ...)

# iterations	
# the number of iterations used in robust fitting, i.e. only if family is "symmetric".
# 
# surface	
# should the fitted surface be computed exactly ("direct") or via interpolation from a k-d tree? Can be abbreviated.
```

#### LOESS predictions for D21
```{r}
# tmp.loess_prediction.D21.df

loess_fit_d21 <- list()
loess_pred_d21 <- list()
gg_t21byage_cluster_trajectories_d21 <- list()
f_t21byage_cluster_trajectories_d21 <- list()
tval_d21 <- list()
lwr_d21 <- list()
upr_d21 <- list()
for ( i in 1:length(ar_p4c_soma_sigt21byage_d21only_clustering_input.split) ){
#for ( i in 1:1 ){
  set.seed(1234)
  ardata.now <- ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]];
  loess_fit_d21[[i]] <- loess(Zlog10_relative_abundance ~ Age + Female,
                          iterations = 10000,
                          data = ardata.now);
  loess_pred_d21[[i]] <- predict(loess_fit_d21[[i]], se = TRUE);
  tval_d21[[i]] <- qt(p = 0.975, df = loess_pred_d21[[i]]$df)
  lwr_d21[[i]] <- loess_pred_d21[[i]]$fit - tval_d21[[i]]*loess_pred_d21[[i]]$se.fit
  upr_d21[[i]] <- loess_pred_d21[[i]]$fit + tval_d21[[i]]*loess_pred_d21[[i]]$se.fit
  ord_d21 <- order(ardata.now$Age)
  f_t21byage_cluster_trajectories_d21[[i]] <- ardata.now %>%
    mutate(ci95_ll = lwr_d21[[i]][ord_d21], 
           ci95_ul = upr_d21[[i]][ord_d21], 
           predicted_Zlog10_relative_abundance = loess_pred_d21[[i]]$fit[ord_d21], 
           Age = ardata.now$Age[ord_d21], 
           Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_d21])
}
```

#### LOESS predictions for T21
```{r}
loess_fit_t21 <- list()
loess_pred_t21 <- list()
gg_t21byage_cluster_trajectories_t21 <- list()
f_t21byage_cluster_trajectories_t21 <- list()
tval_t21 <- list()
lwr_t21 <- list()
upr_t21 <- list()
for ( i in 1:length(ar_p4c_soma_sigt21byage_t21only_clustering_input.split) ){
#for ( i in 1:1 ){
  set.seed(1234)
  ardata.now <- ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[i]];
  loess_fit_t21[[i]] <- loess(Zlog10_relative_abundance ~ Age + Female,
                          iterations = 10000,
                          data = ardata.now);
  loess_pred_t21[[i]] <- predict(loess_fit_t21[[i]], se = TRUE);
  tval_t21[[i]] <- qt(p = 0.975, df = loess_pred_t21[[i]]$df)
  lwr_t21[[i]] <- loess_pred_t21[[i]]$fit - tval_t21[[i]]*loess_pred_t21[[i]]$se.fit
  upr_t21[[i]] <- loess_pred_t21[[i]]$fit + tval_t21[[i]]*loess_pred_t21[[i]]$se.fit
  ord_t21 <- order(ardata.now$Age)
  f_t21byage_cluster_trajectories_t21[[i]] <- ardata.now %>%
    mutate(ci95_ll = lwr_t21[[i]][ord_t21], 
           ci95_ul = upr_t21[[i]][ord_t21], 
           predicted_Zlog10_relative_abundance = loess_pred_t21[[i]]$fit[ord_t21], 
           Age = ardata.now$Age[ord_t21], 
           Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_t21])
}
```

#### LOESS predictions for D21 females (D21 XX)
```{r}
# loess_fit_d21xx <- list()
# loess_pred_d21xx <- list()
# gg_t21byage_cluster_trajectories_d21xx <- list()
# f_t21byage_cluster_trajectories_d21xx <- list()
# tval_d21xx <- list()
# lwr_d21xx <- list()
# upr_d21xx <- list()
# for ( i in 1:length(ar_p4c_soma_sigt21byage_d21only_clustering_input.split) ){
# #for ( i in 1:1 ){
#   set.seed(1234)
#   ardata.now <- ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]] %>%
#     filter(Female == 1);
#   loess_fit_d21xx[[i]] <- loess(Zlog10_relative_abundance ~ Age,
#                           iterations = 10000,
#                           data = ardata.now);
#   loess_pred_d21xx[[i]] <- predict(loess_fit_d21xx[[i]], se = TRUE);
#   tval_d21xx[[i]] <- qt(p = 0.975, df = loess_pred_d21xx[[i]]$df)
#   lwr_d21xx[[i]] <- loess_pred_d21xx[[i]]$fit - tval_d21xx[[i]]*loess_pred_d21xx[[i]]$se.fit
#   upr_d21xx[[i]] <- loess_pred_d21xx[[i]]$fit + tval_d21xx[[i]]*loess_pred_d21xx[[i]]$se.fit
#   ord_d21xx <- order(ardata.now$Age)
#   f_t21byage_cluster_trajectories_d21xx[[i]] <- ardata.now %>%
#     mutate(ci95_ll = lwr_d21xx[[i]][ord_d21xx], 
#            ci95_ul = upr_d21xx[[i]][ord_d21xx], 
#            predicted_Zlog10_relative_abundance = loess_pred_d21xx[[i]]$fit[ord_d21xx], 
#            Age = ardata.now$Age[ord_d21xx], 
#            Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_d21xx])
# }
```

#### LOESS predictions for D21 males (D21 XY)
```{r}
# loess_fit_d21xy <- list()
# loess_pred_d21xy <- list()
# gg_t21byage_cluster_trajectories_d21xy <- list()
# f_t21byage_cluster_trajectories_d21xy <- list()
# tval_d21xy <- list()
# lwr_d21xy <- list()
# upr_d21xy <- list()
# for ( i in 1:length(ar_p4c_soma_sigt21byage_d21only_clustering_input.split) ){
# #for ( i in 1:1 ){
#   set.seed(1234)
#   ardata.now <- ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]] %>%
#     filter(Female == 0);
#   loess_fit_d21xy[[i]] <- loess(Zlog10_relative_abundance ~ Age,
#                           iterations = 10000,
#                           data = ardata.now);
#   loess_pred_d21xy[[i]] <- predict(loess_fit_d21xy[[i]], se = TRUE);
#   tval_d21xy[[i]] <- qt(p = 0.975, df = loess_pred_d21xy[[i]]$df)
#   lwr_d21xy[[i]] <- loess_pred_d21xy[[i]]$fit - tval_d21xy[[i]]*loess_pred_d21xy[[i]]$se.fit
#   upr_d21xy[[i]] <- loess_pred_d21xy[[i]]$fit + tval_d21xy[[i]]*loess_pred_d21xy[[i]]$se.fit
#   ord_d21xy <- order(ardata.now$Age)
#   f_t21byage_cluster_trajectories_d21xy[[i]] <- ardata.now %>%
#     mutate(ci95_ll = lwr_d21xy[[i]][ord_d21xy], 
#            ci95_ul = upr_d21xy[[i]][ord_d21xy], 
#            predicted_Zlog10_relative_abundance = loess_pred_d21xy[[i]]$fit[ord_d21xy], 
#            Age = ardata.now$Age[ord_d21xy], 
#            Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_d21xy])
# }
```

#### LOESS predictions for T21 females (T21 XX)
```{r}
# loess_fit_t21xx <- list()
# loess_pred_t21xx <- list()
# gg_t21byage_cluster_trajectories_t21xx <- list()
# f_t21byage_cluster_trajectories_t21xx <- list()
# tval_t21xx <- list()
# lwr_t21xx <- list()
# upr_t21xx <- list()
# for ( i in 1:length(ar_p4c_soma_sigt21byage_t21only_clustering_input.split) ){
# #for ( i in 1:1 ){
#   set.seed(1234)
#   ardata.now <- ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[i]] %>%
#     filter(Female == 1);
#   loess_fit_t21xx[[i]] <- loess(Zlog10_relative_abundance ~ Age,
#                           iterations = 10000,
#                           data = ardata.now);
#   loess_pred_t21xx[[i]] <- predict(loess_fit_t21xx[[i]], se = TRUE);
#   tval_t21xx[[i]] <- qt(p = 0.975, df = loess_pred_t21xx[[i]]$df)
#   lwr_t21xx[[i]] <- loess_pred_t21xx[[i]]$fit - tval_t21xx[[i]]*loess_pred_t21xx[[i]]$se.fit
#   upr_t21xx[[i]] <- loess_pred_t21xx[[i]]$fit + tval_t21xx[[i]]*loess_pred_t21xx[[i]]$se.fit
#   ord_t21xx <- order(ardata.now$Age)
#   f_t21byage_cluster_trajectories_t21xx[[i]] <- ardata.now %>%
#     mutate(ci95_ll = lwr_t21xx[[i]][ord_t21xx], 
#            ci95_ul = upr_t21xx[[i]][ord_t21xx], 
#            predicted_Zlog10_relative_abundance = loess_pred_t21xx[[i]]$fit[ord_t21xx], 
#            Age = ardata.now$Age[ord_t21xx], 
#            Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_t21xx])
# }
```

#### LOESS predictions for T21 males (T21 XY)
```{r}
# loess_fit_t21xy <- list()
# loess_pred_t21xy <- list()
# gg_t21byage_cluster_trajectories_t21xy <- list()
# f_t21byage_cluster_trajectories_t21xy <- list()
# tval_t21xy <- list()
# lwr_t21xy <- list()
# upr_t21xy <- list()
# for ( i in 1:length(ar_p4c_soma_sigt21byage_t21only_clustering_input.split) ){
# #for ( i in 1:1 ){
#   set.seed(1234)
#   ardata.now <- ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[i]] %>%
#     filter(Female == 0);
#   loess_fit_t21xy[[i]] <- loess(Zlog10_relative_abundance ~ Age,
#                           iterations = 10000,
#                           data = ardata.now);
#   loess_pred_t21xy[[i]] <- predict(loess_fit_t21xy[[i]], se = TRUE);
#   tval_t21xy[[i]] <- qt(p = 0.975, df = loess_pred_t21xy[[i]]$df)
#   lwr_t21xy[[i]] <- loess_pred_t21xy[[i]]$fit - tval_t21xy[[i]]*loess_pred_t21xy[[i]]$se.fit
#   upr_t21xy[[i]] <- loess_pred_t21xy[[i]]$fit + tval_t21xy[[i]]*loess_pred_t21xy[[i]]$se.fit
#   ord_t21xy <- order(ardata.now$Age)
#   f_t21byage_cluster_trajectories_t21xy[[i]] <- ardata.now %>%
#     mutate(ci95_ll = lwr_t21xy[[i]][ord_t21xy], 
#            ci95_ul = upr_t21xy[[i]][ord_t21xy], 
#            predicted_Zlog10_relative_abundance = loess_pred_t21xy[[i]]$fit[ord_t21xy], 
#            Age = ardata.now$Age[ord_t21xy], 
#            Zlog10_relative_abundance = ardata.now$Zlog10_relative_abundance[ord_t21xy])
# }
```

```{r}
ar_sigt21byage_loess_predicted_values <- rbind(f_t21byage_cluster_trajectories_d21 %>% rbindlist(),
                                               f_t21byage_cluster_trajectories_t21 %>% rbindlist()) %>%
                                               # f_t21byage_cluster_trajectories_d21xx %>% rbindlist(),
                                               # f_t21byage_cluster_trajectories_d21xy %>% rbindlist(),
                                               # f_t21byage_cluster_trajectories_t21xx %>% rbindlist(),
                                               # f_t21byage_cluster_trajectories_t21xy %>% rbindlist()) %>%
  mutate(Karyotype_Sex = case_when(T21 == 0 & Female == 1 ~ "D21 XX",
                                   T21 == 0 & Female == 0 ~ "D21 XY",
                                   
                                   T21 == 1 & Female == 1 ~ "T21 XX",
                                   T21 == 1 & Female == 0 ~ "T21 XY"),
         Karyotype_Sex = factor(Karyotype_Sex, levels = c("D21 XX",
                                                          "D21 XY",
                                                          "T21 XX",
                                                          "T21 XY"))) %>%
  dplyr::select(-c(Karyotype_XY));

ar_sigt21byage_loess_predicted_values %>% head();

ar_sigt21byage_loess_predicted_values %>% select(Karyotype, Sex, Karyotype_Sex, T21, Female) %>%
  unique()
```

```{r}
setwd(dir.ardata)
fwrite(ar_sigt21byage_loess_predicted_values, "ar_sigt21byage_loess_predicted_values.csv.gz")
```



# ARCHIVE BELOW FOR NOW (2/23/2025 ~2PM)

```{r}
split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = ci95_ul, 
             ymin = ci95_ll,
             color = Karyotype,
             fill = Karyotype)) + 
  geom_smooth(method = "loess") + # se=T
  geom_point() +
  geom_ribbon(alpha = 0.1, colour=NA);

split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = ci95_ul, 
             ymin = ci95_ll,
             color = Karyotype_Sex,
             fill = Karyotype_Sex)) + 
  geom_smooth(method = "loess") + # 
  geom_point() +
  geom_ribbon(alpha = 0.1, colour=NA);

# Keep
split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = Zlog10_relative_abundance,
             color = Karyotype_Sex,
             fill = Karyotype_Sex)) + 
  geom_smooth(method = "loess", se = T) + #, se=T) + # 
  #geom_point() +
  coord_cartesian(ylim=c(-4, 6.5)) +
  geom_hline(yintercept = 0, linetype = "dashed")
# Keep
split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = Zlog10_relative_abundance,
             color = Karyotype,
             fill = Karyotype)) + 
  geom_smooth(method = "loess", se = T) + #, se=T) + # 
  geom_point() +
  coord_cartesian(ylim=c(-4, 6.5)) +
  geom_hline(yintercept = 0, linetype = "dashed")


split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = ci95_ul, 
             ymin = ci95_ll,
             color = Karyotype,
             fill = Karyotype)) + 
  geom_smooth(method = "loess", se=T) + # 
  geom_point() +
  geom_ribbon(alpha = 0.1, colour=NA);

split(ar_sigt21byage_loess_predicted_values,
      ar_sigt21byage_loess_predicted_values$UniProt_Aptamer_Chr_Cluster)[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = ci95_ul, 
             ymin = ci95_ll,
             color = Karyotype_Sex,
             fill = Karyotype_Sex)) + 
  # Data points
  geom_point(aes(y = predicted_Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_smooth(aes(y = predicted_Zlog10_relative_abundance)) + #, se=T) + 
  # Points from prediction
  geom_point(aes(y = predicted_Zlog10_relative_abundance), stat = "align") + 
  # Confidence interval
  geom_ribbon(alpha = 0.3)
```


#### ARCHIVE BELOW - NOT USING AS OF 2/21/2025

```{r}
rbind(f_t21byage_cluster_trajectories_d21xx[[1]],
      f_t21byage_cluster_trajectories_d21xy[[1]],
      f_t21byage_cluster_trajectories_t21xx[[1]],
      f_t21byage_cluster_trajectories_t21xy[[1]]) %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             #ymax = ci95_ul, 
             #ymin = ci95_ll,
             color = interaction(Karyotype, Sex),
             fill = interaction(Karyotype, Sex))) + 
  geom_smooth(method = "loess", se=T) +
  geom_point();
rbind(f_t21byage_cluster_trajectories_d21xx[[1]],
      f_t21byage_cluster_trajectories_d21xy[[1]],
      f_t21byage_cluster_trajectories_t21xx[[1]],
      f_t21byage_cluster_trajectories_t21xy[[1]]) %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = ci95_ul, 
             ymin = ci95_ll,
             color = interaction(Karyotype, Sex),
             fill = interaction(Karyotype, Sex))) + 
  geom_smooth(method = "loess") + # se=T
  geom_point() +
  geom_ribbon(alpha = 0.1, colour=NA);

f_t21byage_cluster_trajectories_d21xx[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = upr_d21xx, 
             ymin = lwr_d21xx)) + 
  geom_smooth(method = "loess") + # se=T
  geom_point() +
  geom_ribbon(alpha = 0.3);
f_t21byage_cluster_trajectories_d21xy[[1]] %>%
  ggplot(aes(x = Age,
             y = predicted_Zlog10_relative_abundance,
             ymax = upr_d21xx, 
             ymin = lwr_d21xx)) + 
  geom_smooth(method = "loess") + # se=T
  geom_point() +
  geom_ribbon(alpha = 0.3)


f_t21byage_cluster_trajectories_d21xy[[2]]
f_t21byage_cluster_trajectories_t21xy[[2]]
f_t21byage_cluster_trajectories_t21xy[[2]]
```



#### For D21s only, obtain loess predicted values for each aptamer with significant T21*Age interaction
#### ai_loess_predicted_values_d21only
```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.loess_prediction.D21.df <- list()
loess_prediction.D21 <- list()
#plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_sigt21byage_d21only_clustering_input.split) ){
#for ( i in 1:3 ){
  set.seed(1234)
  tmp.loess_prediction.D21.df[[i]] <- ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age + Female,
                                                        # IMPORTANT: ADJUST FOR SEX FOR BETTER CLLUSTERING?
                                                        data = ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[i]]))$fitted)
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  #names(loess_prediction.D21)[[i]] <- names(ar_p4c_soma_d21only_extend_clustering.split)[[i]]
  names(tmp.loess_prediction.D21.df)[[i]] <- names(ar_p4c_soma_sigt21byage_d21only_clustering_input.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()

# old naming:
#loess_prediction.D21.df <- tmp.loess_prediction.D21.df %>%
#  rbindlist();
ai_loess_predicted_values_d21only <- tmp.loess_prediction.D21.df %>%
  rbindlist();

setwd(dir.aidata)
fwrite(ai_loess_predicted_values_d21only, "ai_loess_predicted_values_d21only.csv.gz")

setwd(dir.aidata)
ai_loess_predicted_values_d21only <- fread("ai_loess_predicted_values_d21only.csv.gz")
head(ai_loess_predicted_values_d21only)
dim(ai_loess_predicted_values_d21only) #[1] 87550     49
```

#### For T21s only, obtain loess predicted values for each aptamer with significant T21*Age interaction
#### ai_loess_predicted_values_t21only
```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.loess_prediction.D21.df <- list()
loess_prediction.D21 <- list()
#plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_sigt21byage_t21only_clustering_input.split) ){
#for ( i in 1:3 ){
  set.seed(1234)
  tmp.loess_prediction.D21.df[[i]] <- ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[i]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age + Female,
                                                        # IMPORTANT: ADJUST FOR SEX FOR BETTER CLLUSTERING?
                                                        data = ar_p4c_soma_sigt21byage_t21only_clustering_input.split[[i]]))$fitted)
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  #names(loess_prediction.D21)[[i]] <- names(ar_p4c_soma_t21only_extend_clustering.split)[[i]]
  names(tmp.loess_prediction.D21.df)[[i]] <- names(ar_p4c_soma_sigt21byage_t21only_clustering_input.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_t21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()

# old naming:
#loess_prediction.D21.df <- tmp.loess_prediction.D21.df %>%
#  rbindlist();
ai_loess_predicted_values_t21only <- tmp.loess_prediction.D21.df %>%
  rbindlist();

setwd(dir.aidata)
fwrite(ai_loess_predicted_values_t21only, "ai_loess_predicted_values_t21only.csv.gz")

setwd(dir.aidata)
ai_loess_predicted_values_t21only <- fread("ai_loess_predicted_values_t21only.csv.gz")
head(ai_loess_predicted_values_t21only)
dim(ai_loess_predicted_values_t21only) #[1] 268600     49
```


```{r}
ai_loess_predicted_values_d21only %>% head()
ai_loess_predicted_values_t21only %>% head()
```

#### ar_sigt21byage_loess_predicted_values
```{r}
ar_sigt21byage_loess_predicted_values <- rbind(ai_loess_predicted_values_d21only,
                                   ai_loess_predicted_values_t21only)

ar_sigt21byage_loess_predicted_values %>% head()

setwd(dir.ardata)
fwrite(ar_sigt21byage_loess_predicted_values, "ar_sigt21byage_loess_predicted_values.csv.gz")
```




```{r}
ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = predicted_Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_line(aes(y = predicted_Zlog10_relative_abundance)) + 
  # Points from prediction
  geom_point(aes(y = predicted_Zlog10_relative_abundance)) + 
  # Confidence interval
  geom_ribbon(alpha = 0.3)

ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = predicted_Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_smooth(aes(y = predicted_Zlog10_relative_abundance)) + #, se=T) + 
  # Points from prediction
  geom_point(aes(y = predicted_Zlog10_relative_abundance), stat = "align") + 
  # Confidence interval
  geom_ribbon(alpha = 0.3)

help(geom_ribbon)
```

```{r}
ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_line(aes(y = fit)) + 
  # Points from prediction
  geom_point(aes(y = fit)) + 
  # Confidence interval
  geom_ribbon(alpha = 0.3)

ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_smooth(aes(y = fit)) + 
  # Points from prediction
  geom_point(aes(y = fit)) + 
  # Confidence interval
  geom_ribbon(alpha = 0.3);

  #gg_t21byage_cluster_trajectories[[i]] <- 
ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_line(aes(y = fit)) + 
  # Points from prediction
  geom_point(aes(y = fit)) + 
  # Confidence interval
  geom_ribbon(alpha = 0.3) #+ #, col = "thistle1") + 
  # labs(title = "manual")

  #gg_t21byage_cluster_trajectories[[i]] <- 
ggplot(f_t21byage_cluster_trajectories[[i]],
       aes(x = Age, ymax = upr, ymin = lwr)) + 
  # Data points
  geom_point(aes(y = Zlog10_relative_abundance)) + 
  # Line from prediction
  #geom_line(aes(y = loess_fit[[i]])) + 
  geom_line(aes(y = mean(fit))) + 
  # Points from prediction
  geom_point(aes(y = fit)) + 
  # Confidence interval
  geom_ribbon(alpha = 0.3) #+ #, col = "thistle1") + 
  # labs(title = "manual")


f_t21byage_cluster_trajectories[[i]] %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance,
             ymax = upr, ymin = lwr, color = Karyotype)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(method = "loess", formula = "Zlog10_relative_abundance ~ Age + Female", show.legend = FALSE, se = TRUE) +
  geom_ribbon(alpha = 0.3)


f_t21byage_cluster_trajectories[[i]] %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance,
             ymax = upr, ymin = lwr, color = Karyotype)) + 
  geom_smooth(method = "loess", 
              #formula = "Zlog10_relative_abundance ~ Age + Female",
              se = TRUE)


f_t21byage_cluster_trajectories[[i]] %>%
  ggplot(aes(x = Age, y = fit,
             ymax = upr, ymin = lwr, color = Sex, fill = Sex)) + 
  geom_smooth(method = "loess", 
              #formula = "Zlog10_relative_abundance ~ Age + Female",
              se = TRUE) + 
  geom_point() +
  geom_ribbon(alpha = 0.3)


  geom_smooth(method = "loess",
              formula = "Zlog10_relative_abundance ~ Age + Female",
              #show.legend = FALSE, 
              se = TRUE) + 
  geom_point() +
  geom_ribbon(alpha = 0.3) +
  scale_color_manual(values = RedBlue)
  #labs(x = "Distance", y = "Response") + 
  #theme_classic () +
  #facet_grid(Genotype ~ Treatment)

help(stat_smooth)



length(loess_fit[[i]])

gg_t21byage_cluster_trajectories
```

```{r}
  
data(mtcars)
fit <- loess(mpg ~ hp, data = mtcars)
preds <- predict(fit, se = TRUE)
names(preds)

T <- qt(p = 0.975, df = preds$df)
lwr <- preds$fit - T * preds$se.fit
upr <- preds$fit + T * preds$se.fit

ord <- order(mtcars$hp)
plotData <- data.frame(lwr = lwr[ord], 
                       upr = upr[ord], 
                       fit = preds$fit[ord], 
                       hp = mtcars$hp[ord], 
                       mpg = mtcars$mpg[ord])

p1 <- ggplot(plotData, aes(x = hp, ymax = upr, ymin = lwr)) + 
    #Data points
    geom_point(aes(y = mpg)) + 
    #Line from prediction
    geom_line(aes(y = fit)) + 
    #Points from prediction
    geom_point(aes(y = fit)) + 
    #Confidence interval
    geom_ribbon(alpha = 0.3, col = "thistle1") + 
    labs(title = "manual")
p2 <- ggplot(mtcars, aes(x = hp, y = mpg)) + 
    geom_point() + 
    geom_smooth() + 
    labs(title = "ggplot2")
#Merge plots
library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```

```{r}
ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[10]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age + Female,
                                                        # IMPORTANT: ADJUST FOR SEX FOR BETTER CLLUSTERING?
                                                        data = ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[10]],
                                                        interval = "prediction",
                                                        se.fit = TRUE))$fitted)

predict((loess(Zlog10_relative_abundance ~ Age + Female,
               data = ar_p4c_soma_sigt21byage_d21only_clustering_input.split[[10]])),
        interval = "prediction",
        se.fit = TRUE,
        level = 0.95) %>%
  names()
  as.data.frame()
```

