---
title: "R Notebook"
output: html_notebook
---

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### ar_p4c_soma_sigt21byage_d21only_clustering
#### ar_p4c_soma_sigt21byage_t21only_clustering
```{r}
setwd(dir.ddata)
ar_p4c_soma_sigt21byage_d21only_clustering <- fread("t_p4c_soma_lm_singleaptamer_tidy.csv.gz") %>%
  filter(PadjBH<0.10 & term == "T21:Age") %>%
  left_join(fread(paste0(dir.ardata, "/ar_p4c_soma_d21only.csv.gz")),
            by = c("UniProt_Aptamer_Chr_Cluster", "UniProt", "Aptamer", "Cluster"))

setwd(dir.ddata)
ar_p4c_soma_sigt21byage_t21only_clustering <- fread("t_p4c_soma_lm_singleaptamer_tidy.csv.gz") %>%
  filter(PadjBH<0.10 & term == "T21:Age") %>%
  left_join(fread(paste0(dir.ardata, "/ar_p4c_soma_t21only.csv.gz")),
            by = c("UniProt_Aptamer_Chr_Cluster", "UniProt", "Aptamer", "Cluster"))

setwd(dir.ardata)
fwrite(ar_p4c_soma_sigt21byage_d21only_clustering, "ar_p4c_soma_sigt21byage_d21only_clustering.csv")
fwrite(ar_p4c_soma_sigt21byage_t21only_clustering, "ar_p4c_soma_sigt21byage_t21only_clustering.csv")
```

#### Read ardata for clustering among D21s of protein trajectories with significant T21*Age interaction
```{r}
setwd(dir.ardata)
ar_p4c_soma_sigd21byage_d21only_clustering <- fread("ar_p4c_soma_sigt21byage_d21only_clustering.csv")
#ar_p4c_soma_sigt21byage_t21only_clustering <- fread("ar_p4c_soma_sigt21byage_t21only_clustering.csv")

# ar_p4c_soma_sigd21byage_d21only_clustering

#rm(ar_p4c_soma_sigt21byage_t21only_clustering); gc()
```


```{r}
ar_p4c_soma_sigd21byage_d21only_clustering.split <- ar_p4c_soma_sigd21byage_d21only_clustering %>%
  split(., .$UniProt_Aptamer_Chr_Cluster);

ar_p4c_soma_sigd21byage_d21only_clustering.split[[1]]
```

```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.predicted_values.D21 <- list()
plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_sigd21byage_d21only_clustering.split) ){
  set.seed(1234)
  loess_newclusters.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age + Female,
                          data = ar_p4c_soma_sigd21byage_d21only_clustering.split[[i]])
  
  tmp.predicted_values.D21[[i]] <- cbind(ar_p4c_soma_sigd21byage_d21only_clustering.split[[i]]$LabID,
                                        loess_newclusters.D21[[i]]$fitted,
                                        ar_p4c_soma_sigd21byage_d21only_clustering.split[[i]]$Zlog10_relative_abundance,
                                        ar_p4c_soma_sigd21byage_d21only_clustering.split[[i]]$Age,
                                        as.character(ar_p4c_soma_sigd21byage_d21only_clustering.split[[i]]$Female)) %>%
    as.data.frame() %>%
    `colnames<-`(c("LabID", "predicted_Zlog10_relative_abundance", "Zlog10_relative_abundance", "Age", "Female")) %>%
    mutate(Age = as.numeric(Age),
           predicted_Zlog10_relative_abundance = as.numeric(predicted_Zlog10_relative_abundance),
           Zlog10_relative_abundance = as.numeric(Zlog10_relative_abundance) #,
           #Sex = factor(Sex, levels = c("Male", "Female"))
           ) %>%
    arrange(Age) %>%
    mutate(UniProt_Aptamer_Chr_Cluster = names(ar_p4c_soma_sigd21byage_d21only_clustering.split)[[i]]) %>%
    dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, Age,
                  #Sex,
                  Female,
                  predicted_Zlog10_relative_abundance, Zlog10_relative_abundance) %>%
    separate(UniProt_Aptamer_Chr_Cluster, into = c("UniProt", "Other"), sep = "[|]", extra = "merge", remove = FALSE) %>%
    separate(Other, into = c("Aptamer", "Other"), sep = " [(]", extra = "merge", remove = TRUE) %>%
    separate(Other, into = c("Chromosome", "Lehallier2019_ClusterID"), sep = "[,] ", extra = "merge", remove = TRUE) %>%
    mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID),
           Lehallier2019_ClusterID = trimws(Lehallier2019_ClusterID),
           Chromosome = trimws(Chromosome),
           UniProt = trimws(UniProt),
           Aptamer = trimws(Aptamer)) %>%
  separate(UniProt_Aptamer_Chr_Cluster, into = c("UniProt", "Other"), sep = "[|]", extra = "merge", remove = FALSE) %>%
  separate(Other, into = c("Aptamer", "Other"), sep = " [(]", extra = "merge", remove = TRUE) %>%
  separate(Other, into = c("Chromosome", "Lehallier2019_ClusterID"), sep = "[,] ", extra = "merge", remove = TRUE) %>%
  mutate(Lehallier2019_ClusterID = gsub("[)]", "", Lehallier2019_ClusterID),
         Lehallier2019_ClusterID = trimws(Lehallier2019_ClusterID),
         Chromosome = trimws(Chromosome),
         UniProt = trimws(UniProt),
         Aptamer = trimws(Aptamer))
}
#rm(loess_newclusters.D21, ar_p4c_soma_d21only.split); gc()

ai_predicted_values_d21only <- tmp.predicted_values.D21 %>%
  rbindlist() %>%
#  mutate(Female = case_when(Sex == "Male" ~ 0,
#                            Sex == "Female" ~ 1,
#                            .default = NA)) %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster,
                Lehallier2019_ClusterID,
                LabID, Age,
                #Sex,
                Female,
                predicted_Zlog10_relative_abundance,
                Zlog10_relative_abundance)

# ai_predicted_values_d21only
```

#### ai_predicted_values_d21only.split
```{R}
ai_predicted_values_d21only.split <- ai_predicted_values_d21only %>%
  split(., .$UniProt_Aptamer_Chr_Cluster)

# length(ai_predicted_values_d21only.split) #[1] 4640
```

```{r}
# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
tmp.loess_prediction.D21.df <- list()
loess_prediction.D21 <- list()
#plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ai_predicted_values_d21only.split) ){
#for ( i in 1:3 ){
  set.seed(1234)

  #loess_prediction.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
  #                                    data = ar_p4c_soma_d21only_extend_clustering.split[[i]])

  tmp.loess_prediction.D21.df[[i]] <- ai_predicted_values_d21only.split[[i]] %>%
    mutate(predicted_Zlog10_relative_abundance = (loess(Zlog10_relative_abundance ~ Age + Female, # IMPORTANT: ADJUST FOR SEX FOR BETTER CLLUSTERING?
                                                        data = ai_predicted_values_d21only.split[[i]]))$fitted)
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  #names(loess_prediction.D21)[[i]] <- names(ar_p4c_soma_d21only_extend_clustering.split)[[i]]
  names(tmp.loess_prediction.D21.df)[[i]] <- names(ai_predicted_values_d21only.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()

loess_prediction.D21.df <- tmp.loess_prediction.D21.df %>%
  rbindlist();
#rm(tmp.loess_prediction.D21.df); gc();

# loess_prediction.D21 %>% lapply(class) %>% unique()

loess_prediction.D21.matrix <- tmp.loess_prediction.D21.df %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, predicted_Zlog10_relative_abundance) %>%
  unique() %>%
  spread(key = LabID, value = predicted_Zlog10_relative_abundance) %>%
  column_to_rownames(var = "UniProt_Aptamer_Chr_Cluster") %>%
  as.matrix();
#predicted.D21.matrix %>% as.data.frame()

# Dissimilarity matrix:
set.seed(1234)
d_newclusters.D21s <- dist(loess_prediction.D21.matrix,
                           method = "euclidean");
# Hierarchical clustering using Complete Linkage:
set.seed(1234)
hc1_newclusters.D21s <- hclust(d_newclusters.D21s,
                               method = "complete");
```

```{r}
# # Cut tree into 6 groups
# for ( i in 1:16 ) {
#   #k.now <- i;
#   sub_grp6_newclusters.D21s <- cutree(hc1_newclusters.D21s, k = i)
#   
#   setwd(dir.output)
#   pdf(paste0("gg_p4c_soma_sigt21byage_hctree_k", i, ".pdf"))
#   plot(hc1_newclusters.D21s, cex = 0.1);
#   rect.hclust(hc1_newclusters.D21s, k = i, border = 1:i);
#   dev.off();
# }
# 

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 4, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 8, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 9, border = 1:9);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 10, border = 1:10);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 11, border = 1:11);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 12, border = 1:12);

k_hc1_newclusters.D21s <- 8
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s, border = 1:k_hc1_newclusters.D21s);
```

#### ar_sigt21byage_d21clustered
```{r}
k_hc1_newclusters.D21s <- 8

ar_sigt21byage_d21clustered <- cutree(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s) %>%
    as.data.frame() %>%
    `colnames<-`("T21-by-Age Cluster") %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
    mutate(k = 8) %>%
    ungroup() %>%
  left_join( fread(paste0(dir.ardata, "/ar_p4c_soma.csv.gz")), by = "UniProt_Aptamer_Chr_Cluster" )

head(ar_sigt21byage_d21clustered)

# ar_sigt21byage_d21clustered
```

#### f_cca_barplot_n
#### gg_cca_barplot_n
```{r}
f_cca_barplot_n <- ar_sigt21byage_d21clustered %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster,
                `T21-by-Age Cluster`,
                Lehallier2019_ClusterID) %>%
  unique() %>%
  `rownames<-`(NULL) %>%
  mutate(Lehallier2019_Cluster = case_when(is.na(Lehallier2019_ClusterID)~"None",
                                           .default = as.character(Lehallier2019_ClusterID))) %>%
  dplyr::select(-c(Lehallier2019_ClusterID)) %>%
  group_by(`T21-by-Age Cluster`, Lehallier2019_Cluster) %>%
  summarise(N_Aptamers = n()) %>%
  ungroup() %>%
  group_by(`T21-by-Age Cluster`) %>%
  mutate(N_Total = sum(N_Aptamers)) %>%
  ungroup() %>%
  mutate(Proportion = N_Aptamers/N_Total) %>%
  dplyr::rename(`Proteomic Clock Cluster` = Lehallier2019_Cluster);

setwd(dir.ddata)
fwrite(f_cca_barplot_n, "f_cca_barplot_n.csv")

gg_cca_barplot_n <- f_cca_barplot_n %>%
  ggplot(aes(x = `T21-by-Age Cluster`, y = N_Aptamers,
             color = `Proteomic Clock Cluster`,
             fill = `Proteomic Clock Cluster`,
             width = 0.6)) + 
  geom_bar(position = "dodge", stat = "identity") +
  ylab("Aptamers (N)") +
  theme(aspect.ratio = 1.0,
        legend.position = "right",
        plot.caption = element_text(hjust = 0)) +
  labs(caption = "*As published in Lehallier et al., 2019.") +
  ggtitle("Correspondence of T21-by-Age Clusters with Proteomic Clock Clusters*") +
  # set colors
  scale_colour_manual(values = c("1" = P4C_extended_cluster_colors[[1]],
                                "2" = P4C_extended_cluster_colors[[2]], 
                                "3" = P4C_extended_cluster_colors[[3]], 
                                "4" = P4C_extended_cluster_colors[[4]], 
                                "5" = P4C_extended_cluster_colors[[5]], 
                                "6" = P4C_extended_cluster_colors[[6]], 
                                "7" = P4C_extended_cluster_colors[[7]], 
                                "8" = P4C_extended_cluster_colors[[8]], 
                                "None" = GrayBlue[[1]]))  +
  # set colors
  scale_fill_manual(values = c("1" = P4C_extended_cluster_colors[[1]],
                                "2" = P4C_extended_cluster_colors[[2]], 
                                "3" = P4C_extended_cluster_colors[[3]], 
                                "4" = P4C_extended_cluster_colors[[4]], 
                                "5" = P4C_extended_cluster_colors[[5]], 
                                "6" = P4C_extended_cluster_colors[[6]], 
                                "7" = P4C_extended_cluster_colors[[7]], 
                                "8" = P4C_extended_cluster_colors[[8]], 
                                "None" = GrayBlue[[1]]));

save_as <- "f_cca_barplot_n"
setwd(dir.output)
ggsave(gg_cca_barplot_n,
       filename = paste0(save_as, ".png"),
       width = 11, height = 11, units = "in");

f_cca_barplot_n
gg_cca_barplot_n
```

#### f_cca_barplot_proportion
```{r}
f_cca_barplot_proportion <- ar_sigt21byage_d21clustered %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster,
                `T21-by-Age Cluster`,
                Lehallier2019_ClusterID) %>%
  unique() %>%
  `rownames<-`(NULL) %>%
  mutate(Lehallier2019_Cluster = case_when(is.na(Lehallier2019_ClusterID)~"None",
                                           .default = as.character(Lehallier2019_ClusterID))) %>%
  dplyr::select(-c(Lehallier2019_ClusterID)) %>%
  group_by(`T21-by-Age Cluster`, Lehallier2019_Cluster) %>%
  summarise(N_Aptamers = n()) %>%
  ungroup() %>%
  group_by(`T21-by-Age Cluster`) %>%
  mutate(N_Total = sum(N_Aptamers),
         Proportion = N_Aptamers/N_Total,
         checksum = sum(Proportion)) %>%
  ungroup() %>%
  mutate(Proportion = N_Aptamers/N_Total) %>%
  dplyr::rename(`Proteomic Clock Cluster` = Lehallier2019_Cluster) %>%
  dplyr::select(-c(checksum));

setwd(dir.ddata)
fwrite(f_cca_barplot_proportion, "f_cca_barplot_proportion.csv");

gg_cca_barplot_proportion <- f_cca_barplot_proportion %>%
  ggplot(aes(x = `T21-by-Age Cluster`, y = Proportion,
             color = `Proteomic Clock Cluster`,
             fill = `Proteomic Clock Cluster`)) + 
  geom_bar(position = "stack", stat = "identity") +
  ylab("Aptamers (Proportion)") +
  theme(aspect.ratio = 1.0,
        legend.position = "right",
        plot.caption = element_text(hjust = 0)) +
  labs(caption = "*As published in Lehallier et al., 2019.") +
  ggtitle("Correspondence of T21-by-Age Clusters with Proteomic Clock Clusters*") +
  # set colors
  scale_colour_manual(values = c("1" = P4C_extended_cluster_colors[[1]],
                                "2" = P4C_extended_cluster_colors[[2]], 
                                "3" = P4C_extended_cluster_colors[[3]], 
                                "4" = P4C_extended_cluster_colors[[4]], 
                                "5" = P4C_extended_cluster_colors[[5]], 
                                "6" = P4C_extended_cluster_colors[[6]], 
                                "7" = P4C_extended_cluster_colors[[7]], 
                                "8" = P4C_extended_cluster_colors[[8]], 
                                "None" = GrayBlue[[1]]))  +
  # set colors
  scale_fill_manual(values = c("1" = P4C_extended_cluster_colors[[1]],
                                "2" = P4C_extended_cluster_colors[[2]], 
                                "3" = P4C_extended_cluster_colors[[3]], 
                                "4" = P4C_extended_cluster_colors[[4]], 
                                "5" = P4C_extended_cluster_colors[[5]], 
                                "6" = P4C_extended_cluster_colors[[6]], 
                                "7" = P4C_extended_cluster_colors[[7]], 
                                "8" = P4C_extended_cluster_colors[[8]], 
                                "None" = GrayBlue[[1]])) 

save_as <- "f_cca_barplot_proportion"
setwd(dir.output)
ggsave(gg_cca_barplot_proportion,
       filename = paste0(save_as, ".png"),
       width = 11, height = 11, units = "in");

f_cca_barplot_proportion
gg_cca_barplot_proportion
```

```{r}
# `T21-by-Age Cluster`

n_aptamers <- "_" # stand-in
set.seed(1234)
gg_sigt21byage_d21clustered_bykaryotype <- ar_sigt21byage_d21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", `T21-by-Age Cluster`, "\n(N=", n_aptamers, " with FDR<0.10)")) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype, fill = Karyotype)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = RedBlue) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age Cluster`, ncol = 4)

save_as <- "gg_sigt21byage_d21clustered_bykaryotype"
setwd(dir.output)
ggsave(gg_sigt21byage_d21clustered_bykaryotype,
       filename = paste0(save_as, ".png"),
       width = 11, height = 11, units = "in")
```


```{r}
ar_sigt21byage_d21clustered %>% select(contains("Karyotype")) %>% colnames()
ar_sigt21byage_d21clustered %>% select(Karyotype_Sex) %>% head()
ar_sigt21byage_d21clustered$Karyotype_Sex %>% unique()
#[1] "T21 XX" "T21 XY" "D21 XY" "D21 XX"

gg_sigt21byage_d21clustered_bykaryotypesex <- ar_sigt21byage_d21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", `T21-by-Age Cluster`, "\n(N=", n_aptamers, " with FDR<0.10)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age Cluster`, ncol = 4)

save_as <- "gg_sigt21byage_d21clustered_bykaryotypesex"
setwd(dir.output)
ggsave(gg_sigt21byage_d21clustered_bykaryotypesex,
       filename = paste0(save_as, ".png"),
       width = 11, height = 11, units = "in")
```

