---
title: "R Notebook"
output: html_notebook
---

# KEEP SCRIPT

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata for aging trajectory plots
```{r}
setwd(dir.ardata)
ar_sigt21byage_d21t21clustered <- fread("ar_sigt21byage_d21t21clustered.csv.gz")

ar_nonsigt21byage_d21t21clustered %>% dplyr::select(nonsig_T21byAge_Cluster) %>% unique()
setwd(dir.ardata)
ar_nonsigt21byage_d21t21clustered <- fread("ar_nonsigt21byage_d21t21clustered.csv.gz") %>%
  mutate(nonsig_T21byAge_Cluster = stringr::str_to_title(nonsig_T21byAge_Cluster),
         nonsig_T21byAge_Cluster = factor(nonsig_T21byAge_Cluster,
                                          levels = c("Alpha", "Beta", "Gamma", "Delta", "Epsilon", "Zeta", "Eta")));

#ar_sigt21byage_d21clustered.split %>% head()

ar_sigt21byage_d21t21clustered %>%
  dplyr::select(T21byAge_Cluster, T21byAge_SuperCluster, N_Cluster_Aptamers, N_SuperCluster_Aptamers) %>%
  unique()

ar_nonsigt21byage_d21t21clustered %>%
  dplyr::select(nonsig_T21byAge_Cluster, nonsig_N_Cluster_Aptamers) %>%
  unique()
```

#### f_sigt21byage_d21t21clustered
```{r}
ar_sigt21byage_d21t21clustered %>%
  dplyr::select(T21byAge_Cluster, T21byAge_SuperCluster, N_Cluster_Aptamers, N_SuperCluster_Aptamers) %>%
  unique()

ar_nonsigt21byage_d21t21clustered %>%
  dplyr::select(nonsig_T21byAge_Cluster, nonsig_N_Cluster_Aptamers) %>%
  unique()

f_sigt21byage_d21t21clustered <- ar_sigt21byage_d21t21clustered %>%
  mutate(gg_cluster_title = paste0("T21-by-Age Cluster ", T21byAge_Cluster),
         gg_cluster_subtitle = paste0("(N=", N_Cluster_Aptamers, " Aptamers with FDR<0.05)"),
         gg_supercluster_title = paste0("T21-by-Age SuperCluster ", T21byAge_SuperCluster),
         gg_supercluster_subtitle = paste0("(N=", N_SuperCluster_Aptamers, " Aptamers with FDR<0.05)")) %>%
  ungroup() %>%
  ungroup();

f_sigt21byage_d21t21clustered %>%
  dplyr::select(starts_with("T21byAge_"),
                starts_with("N_"),
                starts_with("gg_")) %>%
  unique()

setwd(dir.ddata)
fwrite(f_sigt21byage_d21t21clustered,
       "f_sigt21byage_d21t21clustered.csv.gz");
```

#### f_nonsigt21byage_d21t21clustered
```{r}
setwd(dir.ardata)
f_nonsigt21byage_d21t21clustered <- fread("ar_nonsigt21byage_d21t21clustered.csv.gz") %>%
  mutate(gg_cluster_title = paste0("Non-sig. T21-by-Age Cluster ", nonsig_T21byAge_Cluster),
         gg_cluster_subtitle = paste0("(N=", nonsig_N_Cluster_Aptamers, " Aptamers with FDR>0.05)"));

f_nonsigt21byage_d21t21clustered %>% dplyr::select(nonsig_T21byAge_Cluster) %>% unique()

setwd(dir.ddata)
fwrite(f_nonsigt21byage_d21t21clustered,
       "f_nonsigt21byage_d21t21clustered.csv.gz")
```

```{r}
f_sigt21byage_d21t21clustered.splitCluster <- f_sigt21byage_d21t21clustered %>%
  split(., .$T21byAge_Cluster)

f_sigt21byage_d21t21clustered.splitSuperCluster <- f_sigt21byage_d21t21clustered %>%
  split(., .$T21byAge_SuperCluster)

rm(f_nonsigt21byage_d21t21clustered); gc()
setwd(dir.ddata)
f_nonsigt21byage_d21t21clustered.split <- fread("f_nonsigt21byage_d21t21clustered.csv.gz") %>% 
  split(., .$nonsig_T21byAge_Cluster)
length(f_nonsigt21byage_d21t21clustered.split)

f_nonsigt21byage_d21t21clustered.split %>% lapply(head)
```

### Aging trajectory plots for sigt21byage_cluster (one file per cluster)

#### gg_trajectory_bykt_sigt21byage_cluster
#### gg_trajectory_byktsex_sigt21byage_cluster
```{r}
gg_trajectory_bykt_sigt21byage_cluster <- list()
gg_trajectory_byktsex_sigt21byage_cluster <- list()
for ( i in 1:length(f_sigt21byage_d21t21clustered.splitCluster) ) {
  gg_trajectory_bykt_sigt21byage_cluster[[i]] <- f_sigt21byage_d21t21clustered.splitCluster[[i]] %>%
    ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype, fill = Karyotype)) +
    geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c(RedBlue)) +
    scale_fill_manual(values = c(RedBlue)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_sigt21byage_d21t21clustered.splitCluster[[i]]$gg_cluster_title)),
         subtitle = element_text(unique(f_sigt21byage_d21t21clustered.splitCluster[[i]]$gg_cluster_subtitle)));
  save_as <- paste0("f_trajectory_bykt_sigt21byage_cluster_",
                    names(f_sigt21byage_d21t21clustered.splitCluster)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_bykt_sigt21byage_cluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")

  gg_trajectory_byktsex_sigt21byage_cluster[[i]] <- f_sigt21byage_d21t21clustered.splitCluster[[i]] %>%
    ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
    geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    #facet_wrap(~T21byAge_Cluster, ncol=4) +
    #ggtitle(paste0("T21-by-Age Cluster ", names(f_sigt21byage_d21t21clustered.splitCluster)[[i]])) +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_sigt21byage_d21t21clustered.splitCluster[[i]]$gg_cluster_title)),
         subtitle = element_text(unique(f_sigt21byage_d21t21clustered.splitCluster[[i]]$gg_cluster_subtitle)));
  save_as <- paste0("f_trajectory_byktsex_sigt21byage_cluster_",
                    names(f_sigt21byage_d21t21clustered.splitCluster)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_byktsex_sigt21byage_cluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")
}
```

### Aging trajectory plots for sigt21byage_cluster (facet by cluster)

#### gg_trajectory_byktsex_sigt21byage_cluster_facet
#### gg_trajectory_byktsex_sigt21byage_cluster_facet_freescale
```{r}
# n_aptamers <- "_n_"
gg_trajectory_byktsex_sigt21byage_cluster_facet <- ar_sigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", T21byAge_Cluster, "\n(N=", N_Cluster_Aptamers, " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 12)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age Cluster`, ncol = 5);

save_as <- "f_trajectory_byktsex_sigt21byage_cluster_facet";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_sigt21byage_cluster_facet,
       filename = paste0(save_as, ".png"),
       width = 13, height = 13, units = "in")

gg_trajectory_byktsex_sigt21byage_cluster_facet_freescale <- ar_sigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", T21byAge_Cluster, "\n(N=", N_Cluster_Aptamers, " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 10.5)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age Cluster`, ncol = 5, scales = "free");

save_as <- "f_trajectory_byktsex_sigt21byage_cluster_facet_freescale";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_sigt21byage_cluster_facet_freescale,
       filename = paste0(save_as, ".png"),
       width = 14, height = 14, units = "in")
```

### Aging trajectory plots for sigt21byage_supercluster (one file per supercluster)

#### gg_trajectory_bykt_sigt21byage_supercluster
#### gg_trajectory_byktsex_sigt21byage_supercluster
```{r}
#rm(f_sigt21byage_d21t21clustered.splitCluster); gc()

gg_trajectory_bykt_sigt21byage_supercluster <- list()
gg_trajectory_byktsex_sigt21byage_supercluster <- list()
for ( i in 1:length(f_sigt21byage_d21t21clustered.splitSuperCluster) ) {
  gg_trajectory_bykt_sigt21byage_supercluster[[i]] <-
    f_sigt21byage_d21t21clustered.splitSuperCluster[[i]] %>%
    ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype, fill = Karyotype)) +
    geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c(RedBlue)) +
    scale_fill_manual(values = c(RedBlue)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_sigt21byage_d21t21clustered.splitSuperCluster[[i]]$gg_supercluster_title)),
         subtitle = element_text(unique(f_sigt21byage_d21t21clustered.splitSuperCluster[[i]]$gg_supercluster_subtitle)));
  save_as <- paste0("f_trajectory_bykt_sigt21byage_supercluster_",
                    names(f_sigt21byage_d21t21clustered.splitSuperCluster)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_bykt_sigt21byage_supercluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")

  gg_trajectory_byktsex_sigt21byage_supercluster[[i]] <-
    f_sigt21byage_d21t21clustered.splitSuperCluster[[i]] %>%
    ggplot(aes(x = Age,
               y = Zlog10_relative_abundance, 
               color = Karyotype_Sex, 
               fill = Karyotype_Sex)) +
    geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_sigt21byage_d21t21clustered.splitSuperCluster[[i]]$gg_supercluster_title)),
         subtitle = element_text(unique(f_sigt21byage_d21t21clustered.splitSuperCluster[[i]]$gg_supercluster_subtitle)));
  save_as <- paste0("f_trajectory_byktsex_sigt21byage_supercluster_",
                    names(f_sigt21byage_d21t21clustered.splitSuperCluster)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_byktsex_sigt21byage_supercluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")
}

f_sigt21byage_d21t21clustered.splitSuperCluster %>% rbindlist() %>%
  dplyr::select(contains("title"))
```

### Aging trajectory plots for sigt21byage_supercluster (facet by supercluster)

#### gg_trajectory_byktsex_sigt21byage_supercluster_facet
#### gg_trajectory_byktsex_sigt21byage_supercluster_facet_freescale
```{r}
gg_trajectory_byktsex_sigt21byage_supercluster_facet <- f_sigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age SuperCluster` = paste0("T21-by-Age SuperCluster ",
                                            T21byAge_SuperCluster,
                                            "\n(N=", N_SuperCluster_Aptamers, " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age SuperCluster`, ncol = 3) +
  ggtitle("Differential Aging in Down syndrome (DA-DS) Clusters");

save_as <- "f_trajectory_byktsex_sigt21byage_supercluster_facet";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_sigt21byage_supercluster_facet,
       filename = paste0(save_as, ".png"),
       width = 15, height = 15, units = "in")

gg_trajectory_byktsex_sigt21byage_supercluster_facet_freescale <- ar_sigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `T21-by-Age SuperCluster` = paste0("T21-by-Age SuperCluster ",
                                            T21byAge_SuperCluster,
                                            "\n(N=", N_SuperCluster_Aptamers, " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`T21-by-Age SuperCluster`, ncol = 3, scales = "free") +
  ggtitle("Differential Aging in Down syndrome (DA-DS) Clusters");

save_as <- "f_trajectory_byktsex_sigt21byage_supercluster_facet_freescale";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_sigt21byage_supercluster_facet_freescale,
       filename = paste0(save_as, ".png"),
       width = 15, height = 15, units = "in")
```

### Aging trajectory plots for NONsigt21byage_cluster (one file per cluster)

#### gg_trajectory_bykt_nonsigt21byage_cluster
#### gg_trajectory_byktsex_nonsigt21byage_cluster
```{r}
gg_trajectory_bykt_nonsigt21byage_cluster <- list()
for ( i in 1:length(f_nonsigt21byage_d21t21clustered.split) ) {
  gg_trajectory_bykt_nonsigt21byage_cluster[[i]] <- f_nonsigt21byage_d21t21clustered.split[[i]] %>%
    ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype, fill = Karyotype)) +
    geom_smooth(method = "lm", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c(RedBlue)) +
    scale_fill_manual(values = c(RedBlue)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_nonsigt21byage_d21t21clustered.split[[i]]$gg_cluster_title)),
         subtitle = element_text(unique(f_nonsigt21byage_d21t21clustered.split[[i]]$gg_cluster_subtitle)));
  save_as <- paste0("f_trajectory_bykt_nonsigt21byage_cluster_",
                    names(f_nonsigt21byage_d21t21clustered.split)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_bykt_nonsigt21byage_cluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")
}

gg_trajectory_byktsex_nonsigt21byage_cluster <- list()
for ( i in 1:length(f_nonsigt21byage_d21t21clustered.split) ) {
  gg_trajectory_byktsex_nonsigt21byage_cluster[[i]] <- f_nonsigt21byage_d21t21clustered.split[[i]] %>%
    ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
    geom_smooth(method = "lm", linewidth = 0.75, se = TRUE) +
    scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme(aspect.ratio = 1.0,
        legend.position = "bottom") +
    ylab("Relative abundance (Zlog10)") +
    theme(aspect.ratio = 1.0,
          legend.position = "bottom",
          legend.title = element_blank()) +
    labs(title = element_text(unique(f_nonsigt21byage_d21t21clustered.split[[i]]$gg_cluster_title)),
         subtitle = element_text(unique(f_nonsigt21byage_d21t21clustered.split[[i]]$gg_cluster_subtitle)));
  save_as <- paste0("f_trajectory_byktsex_nonsigt21byage_cluster_",
                    names(f_nonsigt21byage_d21t21clustered.split)[[i]]);
  setwd(dir.output)
  ggsave(gg_trajectory_byktsex_nonsigt21byage_cluster[[i]],
       filename = paste0(save_as, ".png"),
       width = 5, height = 5, units = "in")
}
```

```{r}
f_nonsigt21byage_d21t21clustered.split %>%
  rbindlist() %>%
  select(Karyotype, Sex, T21, Female, contains("Karyotype")) %>%
  unique()
```

```{r}
setwd(dir.ardata)
list.files()
```


### Aging trajectory plots for NONsigt21byage_cluster (facet by cluster)

#### gg_trajectory_byktsex_nonsigt21byage_cluster_facet
#### gg_trajectory_byktsex_nonsigt21byage_cluster_facet_freescale
```{r}
colnames(f_nonsigt21byage_d21t21clustered)
f_nonsigt21byage_d21t21clustered %>%
  dplyr::select(nonsig_T21byAge_Cluster) %>%
  unique()

# n_aptamers <- "_n_"
gg_trajectory_byktsex_nonsigt21byage_cluster_facet <- f_nonsigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `Non-sig. T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", nonsig_T21byAge_Cluster,
                                                "\n(N=", 
                                                nonsig_N_Cluster_Aptamers,
                                                " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 12)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`Non-sig. T21-by-Age Cluster`, ncol = 5);

save_as <- "f_trajectory_byktsex_nonsigt21byage_cluster_facet";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_nonsigt21byage_cluster_facet,
       filename = paste0(save_as, ".png"),
       width = 15, height = 15, units = "in")

gg_trajectory_byktsex_nonsigt21byage_cluster_facet_freescale <- f_nonsigt21byage_d21t21clustered %>%
  mutate(Karyotype = factor(Karyotype, levels = c("D21", "T21")),
         `Non-sig. T21-by-Age Cluster` = paste0("T21-by-Age Cluster ", nonsig_T21byAge_Cluster,
                                                "\n(N=", nonsig_N_Cluster_Aptamers, " Aptamers with FDR<0.05)"),
         Karyotype_Sex = factor(Karyotype_Sex, c("D21 XX", "D21 XY", "T21 XX", "T21 XY"))) %>%
  arrange(Karyotype) %>%
  ggplot(aes(x = Age, y = Zlog10_relative_abundance, color = Karyotype_Sex, fill = Karyotype_Sex)) +
  geom_smooth(method = "loess", linewidth = 0.75, se = TRUE) +
  scale_colour_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  scale_fill_manual(values = c("pink3",
                                 RedBlue[[1]],
                                 "navy",
                                 RedBlue[[2]])) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(aspect.ratio = 1.0,
        legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(size = 12),
        plot.subtitle = element_text(size = 10.5)) +
  ylab("Relative abundance (Zlog10)") +
  facet_wrap(~`Non-sig. T21-by-Age Cluster`, ncol = 5, scales = "free");

save_as <- "f_trajectory_byktsex_nonsigt21byage_cluster_facet_freescale";
setwd(dir.output)
ggsave(gg_trajectory_byktsex_nonsigt21byage_cluster_facet_freescale,
       filename = paste0(save_as, ".png"),
       width = 15, height = 15, units = "in")
```



