---
title: "R Notebook"
output: html_notebook
---

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata
```{r}
setwd(dir.ardata)
ar_p4c_soma_d21only <- fread("ar_p4c_soma_d21only.csv.gz") #%>%
  #filter(Aptamer_in_P4C == 1 & Aptamer_in_Lehallier2019 == 0);

setwd(dir.ardata)
ar_p4c_soma_t21only <- fread("ar_p4c_soma_t21only.csv.gz")

#ar_p4c_soma_d21only
#ar_p4c_soma_t21only
```

```{r}
ar_p4c_soma_d21only %>%
  dplyr::select(Aptamer, Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>%
  unique() %>%
  group_by(Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>% 
  summarise(N = n())

ar_p4c_soma_t21only %>%
  dplyr::select(Aptamer, Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>%
  unique() %>%
  group_by(Aptamer_in_Lehallier2019, Aptamer_in_P4C) %>% 
  summarise(N = n())
```

#### *Within D21s only*, cluster trajectories of all proteins measured in P4C, even if already clustered and reported in Lehallier et al., 2019
```{R}
ar_p4c_soma_d21only.split <- ar_p4c_soma_d21only %>%
  split(., .$UniProt_Aptamer_Chr_Cluster)

# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.D21 <- list()
predicted_newclusters.D21 <- list()
plot_predicted_newclusters.D21 <- list()
for ( i in 1:length(ar_p4c_soma_d21only.split) ){
  set.seed(1234)

  loess_newclusters.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
                          data = ar_p4c_soma_d21only.split[[i]])
  predicted_newclusters.D21[[i]] <- cbind(ar_p4c_soma_d21only.split[[i]]$LabID,
                                        loess_newclusters.D21[[i]]$fitted,
                                        ar_p4c_soma_d21only.split[[i]]$Zlog10_relative_abundance,
                                        ar_p4c_soma_d21only.split[[i]]$Age,
                                        as.character(ar_p4c_soma_d21only.split[[i]]$Sex)) %>%
    as.data.frame() %>%
    `colnames<-`(c("LabID", "predicted_Zlog10_relative_abundance", "Zlog10_relative_abundance", "Age", "Sex")) %>%
    mutate(Age = as.numeric(Age),
           predicted_Zlog10_relative_abundance = as.numeric(predicted_Zlog10_relative_abundance),
           Zlog10_relative_abundance = as.numeric(Zlog10_relative_abundance),
           Sex = factor(Sex, levels = c("Male", "Female"))) %>%
    arrange(Age) %>%
    mutate(UniProt_Aptamer_Chr_Cluster = names(ar_p4c_soma_d21only.split)[[i]]) %>%
    dplyr::select(UniProt_Aptamer_Chr_Cluster, everything())
  
  # plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
  #   ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
  #   geom_point(color = RedBlue[[1]], alpha = 0.5) +
  #   geom_line(color = RedBlue[[1]]) +
  #   ggtitle(names(predicted_newclusters.D21)[[i]])
  
  names(loess_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
  names(predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
  #names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
}
#predicted_newclusters.D21 %>% head()
#plot_predicted_newclusters.D21 %>% head()

set.seed(1234)
predicted_newclusters.D21.matrix <- predicted_newclusters.D21 %>%
  rbindlist() %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID, predicted_Zlog10_relative_abundance) %>%
  unique() %>%
  spread(key = LabID, value = predicted_Zlog10_relative_abundance) %>%
  column_to_rownames(var = "UniProt_Aptamer_Chr_Cluster") %>%
  as.matrix();
#predicted.D21.matrix %>% as.data.frame()

# Dissimilarity matrix:
d_newclusters.D21s <- dist(predicted_newclusters.D21.matrix,
                           method = "euclidean");
# Hierarchical clustering using Complete Linkage:
hc1_newclusters.D21s <- hclust(d_newclusters.D21s,
                               method = "complete");
# Plot the obtained dendrogram:
k.now <- 8;
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k.now, border = 1:k.now);
k.now <- 11;
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k.now, border = 1:k.now);
k.now <- 12;
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k.now, border = 1:k.now);

# Cut tree into 6 groups
k.now <- 12;
sub_grp6_newclusters.D21s <- cutree(hc1_newclusters.D21s, k = k.now)
# sub_grp6_newclusters.D21s

# Add the the cluster each observation belongs to to our original data.
length(sub_grp6_newclusters.D21s) #[1] 4632
predicted_newclusters.D21.matrix %>% as.data.frame() %>% nrow() #[1] 4632

ar_p4c_soma_d21only_aging_clusters <- predicted_newclusters.D21.matrix %>%
  as.data.frame() %>%
  #mutate(P4C_Clock_Cluster = 8 + sub_grp6_newclusters.D21s) %>% # Assign cluster numbers 9-16 since Lehallier 2019 defined clusters 1-8. # First option tried...
    mutate(P4C_Clock_Cluster = sub_grp6_newclusters.D21s) %>% # New option from clustering ALL measured proteins.
  dplyr::select(P4C_Clock_Cluster, everything()) %>%
  rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
  gather(key = "LabID", value = "predicted_Zlog10_relative_abundance", starts_with("HTP")) %>%
  ungroup() %>%
  mutate(LabID = trimws(LabID),
         UniProt_Aptamer_Chr_Cluster = trimws(UniProt_Aptamer_Chr_Cluster)) %>%
  left_join(( (fread(paste0(dir.ardata, "/ar_p4c_soma_d21only.csv.gz"))) %>%
              dplyr::select(LabID, Age, Sex, UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID) %>%
               unique() %>%
               filter(!is.na(Age)) %>%
               ungroup() %>%
               mutate(LabID = trimws(LabID),
                      UniProt_Aptamer_Chr_Cluster = trimws(UniProt_Aptamer_Chr_Cluster))),
            by = c("LabID", "UniProt_Aptamer_Chr_Cluster")) %>%
  dplyr::select(LabID, Age, Sex, UniProt_Aptamer_Chr_Cluster, 
         Lehallier2019_ClusterID, P4C_Clock_Cluster,
         predicted_Zlog10_relative_abundance) %>%
  full_join( (fread(paste0(dir.ardata, "/ar_p4c_soma_d21only.csv.gz")) %>%
     #filter(Aptamer_in_P4C == 1 & Aptamer_in_Lehallier2019 == 1) %>%
     dplyr::select(LabID, Age, Sex, UniProt_Aptamer_Chr_Cluster,
                   Lehallier2019_ClusterID, everything())),
     by = c("LabID", "Age", "Sex", "UniProt_Aptamer_Chr_Cluster", "Lehallier2019_ClusterID")) %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID,
                P4C_Clock_Cluster, everything()) %>%
  unique() %>%
  #mutate(Clock_Cluster = case_when(is.na(Lehallier2019_ClusterID) ~ P4C_Clock_Cluster,
  #                                 .default = Lehallier2019_ClusterID)) %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster, Clock_Cluster,
                everything());


# ar_p4c_soma_d21_clock_clusters_extended %>%
#   dplyr::select(UniProt_Aptamer_Chr_Cluster, Lehallier2019_ClusterID, P4C_Clock_Cluster, Clock_Cluster) %>% unique() %>%
#   split(., .$Clock_Cluster)

#ar_p4c_soma_d21_clock_clusters_extended

#setwd(dir.ardata)
#fwrite(ar_p4c_soma_d21_clock_clusters_extended, "ar_p4c_soma_d21_clock_clusters_extended.csv.gz")

setwd(dir.ardata)
fwrite(ar_p4c_soma_d21only_aging_clusters, "ar_p4c_soma_d21only_aging_clusters.csv.gz")
```

#### ar_p4c_soma_d21_clock_clusters_extended
```{r}

```



#### *Within T21s only*, cluster trajectories of all proteins measured in P4C, even if already clustered and reported in Lehallier et al., 2019
```{R}
ar_p4c_soma_t21only.split <- ar_p4c_soma_t21only %>%
  split(., .$UniProt_Aptamer_Chr_Cluster)

# https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
loess_newclusters.T21 <- list()
predicted_newclusters.T21 <- list()
plot_predicted_newclusters.T21 <- list()
for ( i in 1:length(ar_p4c_soma_t21only.split) ){
  set.seed(1234)

  loess_newclusters.T21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
                          data = ar_p4c_soma_t21only.split[[i]])
  predicted_newclusters.T21[[i]] <- cbind(ar_p4c_soma_t21only.split[[i]]$LabID,
                                        loess_newclusters.T21[[i]]$fitted,
                                        ar_p4c_soma_t21only.split[[i]]$Zlog10_relative_abundance,
                                        ar_p4c_soma_t21only.split[[i]]$Age,
                                        as.character(ar_p4c_soma_t21only.split[[i]]$Sex)) %>%
    as.data.frame() %>%
    `colnames<-`(c("LabID", "predicted_Zlog10_relative_abundance", "Zlog10_relative_abundance", "Age", "Sex")) %>%
    mutate(Age = as.numeric(Age),
           predicted_Zlog10_relative_abundance = as.numeric(predicted_Zlog10_relative_abundance),
           Zlog10_relative_abundance = as.numeric(Zlog10_relative_abundance),
           Sex = factor(Sex, levels = c("Male", "Female"))) %>%
    arrange(Age) %>%
    mutate(UniProt_Aptamer_Chr_Cluster = names(ar_p4c_soma_t21only.split)[[i]]) %>%
    dplyr::select(UniProt_Aptamer_Chr_Cluster, everything())
  
  plot_predicted_newclusters.T21[[i]] <- predicted_newclusters.T21[[i]] %>%
    ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
    geom_point(color = RedBlue[[1]], alpha = 0.5) +
    geom_line(color = RedBlue[[1]]) +
    ggtitle(names(predicted_newclusters.T21)[[i]])
  
  names(loess_newclusters.T21)[[i]] <- names(ar_p4c_soma_t21only.split)[[i]]
  names(predicted_newclusters.T21)[[i]] <- names(ar_p4c_soma_t21only.split)[[i]]
  names(plot_predicted_newclusters.T21)[[i]] <- names(ar_p4c_soma_t21only.split)[[i]]
}

predicted_newclusters.T21 %>% head()

plot_predicted_newclusters.T21 %>% head()
```


#### f_p4c_soma_d21_clock_clusters_extended
```{r}
# KEEP:
# set.seed(1234)
# factoextra::fviz_cluster(object = list(data = predicted_newclusters.D21.matrix,
#                                         cluster = cutree(hc1_newclusters.D21s, k = k.now)),
#                          geom = "point",
#                          repel = TRUE);

f_p4c_soma_d21_clock_clusters_extended <- ar_p4c_soma_d21_clock_clusters_extended %>%
  mutate(Clock_Cluster = paste0("Clock Cluster ", Clock_Cluster),
         Clock_Cluster = factor(Clock_Cluster,
                                levels = c(paste0("Clock Cluster ", seq(1:18))))) %>%
  arrange(Clock_Cluster);

# f_p4c_soma_d21_clock_clusters_extended

#### Write ddata to file
setwd(dir.ddata)
fwrite(f_p4c_soma_d21_clock_clusters_extended, "f_p4c_soma_d21_clock_clusters_extended.csv.gz")
```



#### ARCHIVE

#### (JK) Cluster protein trajectories NOT already clustered by Lehallier et al. 2019, within D21s only
```{r}
# setwd(dir.ardata)
# ar_p4c_soma_d21only <- fread("ar_p4c_soma_d21only.csv.gz") %>%
#   filter(Aptamer_in_P4C == 1 & Aptamer_in_Lehallier2019 == 0);
# 
# ar_p4c_soma_d21only.split <- ar_p4c_soma_d21only %>%
#   split(., .$UniProt_Aptamer_Chr_Cluster)
# 
# # https://stackoverflow.com/questions/15337777/fit-a-line-with-loess-in-r
# loess_newclusters.D21 <- list()
# predicted_newclusters.D21 <- list()
# plot_predicted_newclusters.D21 <- list()
# for ( i in 1:length(ar_p4c_soma_d21only.split) ){
#   set.seed(1234)
# 
#   loess_newclusters.D21[[i]] <- loess(Zlog10_relative_abundance ~ Age,
#                           data = ar_p4c_soma_d21only_newclusters.split[[i]])
#   predicted_newclusters.D21[[i]] <- cbind(ar_p4c_soma_d21only.split[[i]]$LabID,
#                                         loess_newclusters.D21[[i]]$fitted,
#                                         ar_p4c_soma_d21only.split[[i]]$Zlog10_relative_abundance,
#                                         ar_p4c_soma_d21only.split[[i]]$Age,
#                                         as.character(ar_p4c_soma_d21only.split[[i]]$Sex)) %>%
#     as.data.frame() %>%
#     `colnames<-`(c("LabID", "predicted_Zlog10_relative_abundance", "Zlog10_relative_abundance", "Age", "Sex")) %>%
#     mutate(Age = as.numeric(Age),
#            predicted_Zlog10_relative_abundance = as.numeric(predicted_Zlog10_relative_abundance),
#            Zlog10_relative_abundance = as.numeric(Zlog10_relative_abundance),
#            Sex = factor(Sex, levels = c("Male", "Female"))) %>%
#     arrange(Age) %>%
#     mutate(UniProt_Aptamer_Chr_Cluster = names(ar_p4c_soma_d21only.split)[[i]]) %>%
#     dplyr::select(UniProt_Aptamer_Chr_Cluster, everything())
#   
#   plot_predicted_newclusters.D21[[i]] <- predicted_newclusters.D21[[i]] %>%
#     ggplot(aes(x = Age, y = predicted_Zlog10_relative_abundance)) +
#     geom_point(color = RedBlue[[1]], alpha = 0.5) +
#     geom_line(color = RedBlue[[1]]) +
#     ggtitle(names(predicted_newclusters.D21)[[i]])
#   
#   names(loess_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
#   names(predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
#   names(plot_predicted_newclusters.D21)[[i]] <- names(ar_p4c_soma_d21only.split)[[i]]
# }
# 
# predicted_newclusters.D21 %>% head()
# 
# plot_predicted_newclusters.D21 %>% head()
```

