---
title: "R Notebook"
output: html_notebook
---

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")

select <- dplyr::select;
filter <- dplyr::filter;
rename <- dplyr::rename;
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
dir.output <- paste0(dir.project, "/output")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1)

analysis_tracker
```

#### Read ardata
```{r}
setwd(dir.ardata)
ar_sigt21byage_loess_predicted_values <- fread("ar_sigt21byage_loess_predicted_values.csv.gz") %>%
  filter(Karyotype == "Control") # IMPORTANT - THIS IS ONE OF THE ONLY CHANGES FROM THE SCRIPT FOR D21T21CLUSTERED

head(ar_sigt21byage_loess_predicted_values)

ar_sigt21byage_loess_predicted_values %>%
  select(#LabID, 
         Karyotype, Sex, Karyotype_Sex) %>%
  unique()
```

#### Prepare matrix of predicted values for input to hierarchical clustering
```{r}
loess_prediction.D21.matrix <- ar_sigt21byage_loess_predicted_values %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, LabID,
                predicted_Zlog10_relative_abundance) %>%
  unique() %>%
  spread(key = LabID, value = predicted_Zlog10_relative_abundance) %>%
  column_to_rownames(var = "UniProt_Aptamer_Chr_Cluster") %>%
  as.matrix();

head(as.data.frame(loess_prediction.D21.matrix))
```

#### Read ardata for clustering among D21s of protein trajectories with significant T21*Age interaction
```{r}
# Dissimilarity matrix:
set.seed(1234)
d_newclusters.D21s <- dist(loess_prediction.D21.matrix,
                           method = "euclidean");
# Hierarchical clustering using Complete Linkage:
set.seed(1234)
hc1_newclusters.D21s <- hclust(d_newclusters.D21s,
                               method = "complete");

```

#### Decide how many T21-by-Age Clusters make sense
```{r}
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 4, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 6, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 7, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 8, border = 1:8);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 9, border = 1:9);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 10, border = 1:10);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 11, border = 1:11);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 12, border = 1:12);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 13, border = 1:13);

plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = 14, border = 1:14);
# After filtering T21-by-Age results by PadjBH<0.05 instead of PadjBH<0.10, we need 14 T21-by-Age Clusters in order to get separation of the two branches of the middle area of the tree.
```

#### Hierarchically cluster the age-dependent trajectories of proteins with significant T21-by-Age interaction into k=14 clusters

After filtering T21-by-Age results by PadjBH<0.05 instead of PadjBH<0.10, we need k=14 T21-by-Age Clusters in order to get separation of the two branches of the middle area of the tree that contains the bulk of significant results.
```{r}
k_hc1_newclusters.D21s <- 8
plot(hc1_newclusters.D21s, cex = 0.1);
rect.hclust(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s, 
            border = 1:k_hc1_newclusters.D21s);
```

#### Use the cuttree function to assign aptamers to k=8 T21-by-Age Clusters ('ar_sigt21byage_d21onlyclustered')
```{r echo=T, results='hide'}
k_hc1_newclusters.D21s <- 8
set.seed(1234)
tmp.ar_sigt21byage_d21onlyclustered <- cutree(hc1_newclusters.D21s, k = k_hc1_newclusters.D21s) %>%
    as.data.frame() %>%
    `colnames<-`("D21_Clock_Cluster") %>%
    rownames_to_column("UniProt_Aptamer_Chr_Cluster") %>%
    mutate(k = k_hc1_newclusters.D21s) %>%
    ungroup() %>%
  left_join(ar_sigt21byage_loess_predicted_values, by = "UniProt_Aptamer_Chr_Cluster") #%>%
  # mutate(D21_Clock_Cluster = case_when(D21_Clock_Cluster == 1 ~ "A",
  #                                     D21_Clock_Cluster == 2 ~ "B",
  #                                     D21_Clock_Cluster == 3 ~ "C",
  #                                     D21_Clock_Cluster == 4 ~ "D",
  #                                     D21_Clock_Cluster == 5 ~ "E",
  #                                     D21_Clock_Cluster == 6 ~ "F",
  #                                     D21_Clock_Cluster == 7 ~ "G",
  #                                     D21_Clock_Cluster == 8 ~ "H"))

ar_sigt21byage_d21onlyclustered <- tmp.ar_sigt21byage_d21onlyclustered %>%
  dplyr::select(D21_Clock_Cluster, UniProt_Aptamer_Chr_Cluster) %>%
  unique() %>%
  group_by(D21_Clock_Cluster) %>%
  summarise(N_Cluster_Aptamers = n()) %>%
  ungroup() %>%
  full_join(tmp.ar_sigt21byage_d21onlyclustered, by = "D21_Clock_Cluster") 

#left_join( fread(paste0(dir.ardata,
  #                        "/ar_sigt21byage_loess_predicted_values.csv.gz")),
  #           by = "UniProt_Aptamer_Chr_Cluster" )

head(tmp.ar_sigt21byage_d21onlyclustered)
```

#### ar_sigt21byage_d21onlyclustered
```{r}
setwd(dir.ardata)
fwrite(ar_sigt21byage_d21onlyclustered, "ar_sigt21byage_d21onlyclustered.csv.gz")
```


```{r}
ar_sigt21byage_d21onlyclustered %>%
  dplyr::select(Aptamer, D21_Clock_Cluster, Lehallier2019_ClusterID, Aptamer_in_P4C, Aptamer_in_Lehallier2019) %>%
  unique() %>%
  dplyr::filter(Aptamer_in_Lehallier2019==1 & Aptamer_in_P4C==1) %>%
  group_by(Aptamer_in_P4C, Aptamer_in_Lehallier2019, Lehallier2019_ClusterID, D21_Clock_Cluster) %>%
  summarise(N = n())
```



```{r}
ar_sigt21byage_d21onlyclustered %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, T21byAge_Cluster) %>%
  unique() %>%
  group_by(T21byAge_Cluster) %>%
  summarise(N = n())

ar_sigt21byage_d21onlyclustered %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, T21byAge_Cluster, Lehallier2019_ClusterID) %>%
  unique() %>%
  group_by(T21byAge_Cluster, Lehallier2019_ClusterID) %>%
  summarise(N = n())
```

```{r}
ar_sigt21byage_aptamer_clusters <- ar_sigt21byage_d21onlyclustered %>%
  dplyr::select(UniProt_Aptamer_Chr_Cluster, T21byAge_Cluster) %>%
  unique()

ar_sigt21byage_aptamer_clusters

ar_sigt21byage_aptamer_clusters %>%
  split(., .$T21byAge_Cluster)
```


