---
title: "R Notebook"
output: html_notebook
---

#### Overview of fitted models
##### Model C: Zlog10_relative_abundance ~ Age + Female (D21s only)
###### Hypotheses tested by model:
- Difference in the mean value of Zlog10_relative_abundance for each 1 year increase in age at blood draw (i.e., the slope for age), among D21s only.
- Difference in the group mean of Zlog10_relative_abundance for female D21s as compared to male D21s.

#### Run startup function
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"

setwd(paste0(dir.project, "/macro"))
source("ui_init.R")
```

#### Define paths to directories
```{r}
dir.project <- "~/Dropbox/ShawJR/2025/dev"
dir.donovan2024_source <- paste0(dir.project, "/rawdata/Donovan2024_syn31481952.5")
dir.indata <- paste0(dir.project, "/indata")
dir.aidata <- paste0(dir.project, "/aidata")
dir.ardata <- paste0(dir.project, "/ardata")
dir.ddata <- paste0(dir.project, "/ddata")
dir.docs <- paste0(dir.project, "/docs")
```

#### Read Analysis Tracker
```{r}
setwd(dir.docs)
analysis_tracker <- read.xlsx("AnalysisTracker.xlsx") %>%
  filter(run_model == 1) %>%
  filter(ddata_name == gsub(".Rmd", "", basename(rstudioapi::getSourceEditorContext()$path)))

analysis_tracker
```

#### Read ardata
```{r}
setwd(dir.ardata)
ardata.split <- fread(paste0(unique(analysis_tracker$ardata_name), ".csv.gz")) %>%
  filter(Aptamer_in_P4C == 1) %>%
  split(., .$UniProt_Aptamer_Chr_Cluster)
# Note: the above two lines of code should be the only revision to ardata before modeling.

ardata.split %>% rbindlist()
```

#### Vectorize and name the model formulas to be fit later in this script
modeld: Zlog10_relative_abundance ~ Age + Female (D21s only)
```{r}
tmp <- analysis_tracker %>%
  dplyr::select(model_formula) %>%
  unique() %>%
  mutate(model_name = paste0("model", row_number())) %>%
  `rownames<-`(NULL) %>%
  mutate(model_name = "modelD");
model_formulas <- tmp$model_formula;
names(model_formulas) <- tmp$model_name;

model_formulas
```

#### Run linear models pre-specified in '/docs/AnalysisTracker.xlsx'
```{r, echo = F, results = F}
options(future.globals.maxSize = 5e10);

model_formula.now <- model_formulas[[1]]
t0 <- Sys.time()
set.seed(1234)
lm_fit <- ardata.split %>%
    purrr::map(~lm(as.formula(model_formula.now), data = .));
lm_vif <- ardata.split %>%
    purrr::map(~car::vif(lm(as.formula(model_formula.now), data = .))) %>%
    lapply(as.data.frame) %>%
    lapply(rownames_to_column, "variable") %>%
    rbindlist(fill = TRUE, idcol = "id") %>%
    mutate(model_formula = model_formula.now) %>%
    dplyr::select(model_formula, everything());
lm_tidy <- ardata.split %>%
    purrr::map(~lm(as.formula(model_formula.now), data = .)) %>%
    purrr::map(.x = .,
         .f = ~broom::tidy(x = .x, conf.int=TRUE, data = .y)) %>%
      bind_rows(.id="UniProt_Aptamer_Chr_Cluster") %>%
    mutate(model_formula = model_formula.now) %>%
    dplyr::select(model_formula, everything()) %>%
    filter(term!="(Intercept)") %>%
    group_by(UniProt_Aptamer_Chr_Cluster, term) %>% # IMPORTANT: MUST GROUP BY BOTH TERM AND APTAMER.
    mutate(PadjBH = p.adjust(p.value, method = "BH")) %>%
    ungroup() %>%
    dplyr::rename(Zlog10FC = estimate, P = p.value) %>%
    dplyr::select(UniProt_Aptamer_Chr_Cluster, term, Zlog10FC, P, PadjBH, everything());
lm_glance <- ardata.split %>%
    purrr::map(~lm(as.formula(model_formula.now), data = .)) %>%
    purrr::map(.x = .,
         .f = ~glance(x = .x, conf.int=TRUE, data = .y)) %>%
      bind_rows(.id="UniProt_Aptamer_Chr_Cluster") %>%
    mutate(model_formula = model_formula.now) %>%
    dplyr::select(model_formula, everything());
lm_augment <- ardata.split %>%
    purrr::map(~lm(as.formula(model_formula.now), data = .)) %>%
    purrr::map(.x = .,
         .f = ~glance(x = .x, conf.int=TRUE, data = .y)) %>%
      bind_rows(.id="UniProt_Aptamer_Chr_Cluster") %>%
    mutate(model_formula = model_formula.now) %>%
    dplyr::select(model_formula, everything());
t1 <- Sys.time();
tdiff<- t1 - t0;
```

```{r}
t_p4c_soma_lm_singleaptamer_t21only_vif <- lm_vif %>%
  dplyr::rename(VIF = `X[[i]]`,
                UniProt_Aptamer_Chr_Cluster = id) %>%
  separate(UniProt_Aptamer_Chr_Cluster,
           into = c("UniProt", "other"),
           sep = "[|]",
           extra = "merge",
           remove = FALSE) %>%
  separate(other,
           into = c("Aptamer", "other"),
           sep = "[()]",
           extra = "merge",
           remove = TRUE) %>%
  separate(other,
           into = c("Chromosome", "Cluster"),
           sep = ",",
           extra = "merge",
           remove = TRUE) %>%
  mutate(Cluster = gsub(")", "", Cluster)) %>%
    arrange(Chromosome) %>%
    mutate(Chromosome = gsub("Chr", "", Chromosome)) %>%
    mutate(Chromosome = case_when(!is.na(as.numeric(Chromosome)) ~ as.numeric(Chromosome),
                                  Chromosome == "X" ~ 23,
                                  Chromosome == "Y" ~ 24,
                                  Chromosome == "X|Y" ~ 25,
                                  Chromosome == "MT" ~ 26,
                                  .default = NA));
t_p4c_soma_lm_singleaptamer_t21only_vif %>%
  group_by(model_formula, UniProt_Aptamer_Chr_Cluster) %>%
  summarise(maxVIF = max(VIF)) %>%
  ungroup() %>%
  arrange(desc(maxVIF));
# All good. No indication of multicollinearity.

t_p4c_soma_lm_singleaptamer_t21only_tidy <- lm_tidy %>%
  separate(UniProt_Aptamer_Chr_Cluster,
           into = c("UniProt", "other"),
           sep = "[|]",
           extra = "merge",
           remove = FALSE) %>%
  separate(other,
           into = c("Aptamer", "other"),
           sep = "[()]",
           extra = "merge",
           remove = TRUE) %>%
  separate(other,
           into = c("Chromosome", "Cluster"),
           sep = ",",
           extra = "merge",
           remove = TRUE) %>%
  mutate(Cluster = gsub(")", "", Cluster)) %>%
    arrange(Chromosome) %>%
    mutate(Chromosome = gsub("Chr", "", Chromosome)) %>%
    mutate(Chromosome = case_when(!is.na(as.numeric(Chromosome)) ~ as.numeric(Chromosome),
                                  Chromosome == "X" ~ 23,
                                  Chromosome == "Y" ~ 24,
                                  Chromosome == "X|Y" ~ 25,
                                  Chromosome == "MT" ~ 26,
                                  .default = NA));

t_p4c_soma_lm_singleaptamer_t21only_glance <- lm_glance %>%
  separate(UniProt_Aptamer_Chr_Cluster,
           into = c("UniProt", "other"),
           sep = "[|]",
           extra = "merge",
           remove = FALSE) %>%
  separate(other,
           into = c("Aptamer", "other"),
           sep = "[()]",
           extra = "merge",
           remove = TRUE) %>%
  separate(other,
           into = c("Chromosome", "Cluster"),
           sep = ",",
           extra = "merge",
           remove = TRUE) %>%
  mutate(Cluster = gsub(")", "", Cluster)) %>%
    arrange(Chromosome) %>%
    mutate(Chromosome = gsub("Chr", "", Chromosome)) %>%
    mutate(Chromosome = case_when(!is.na(as.numeric(Chromosome)) ~ as.numeric(Chromosome),
                                  Chromosome == "X" ~ 23,
                                  Chromosome == "Y" ~ 24,
                                  Chromosome == "X|Y" ~ 25,
                                  Chromosome == "MT" ~ 26,
                                  .default = NA));

t_p4c_soma_lm_singleaptamer_t21only_augment <- lm_augment %>%
  separate(UniProt_Aptamer_Chr_Cluster,
           into = c("UniProt", "other"),
           sep = "[|]",
           extra = "merge",
           remove = FALSE) %>%
  separate(other,
           into = c("Aptamer", "other"),
           sep = "[()]",
           extra = "merge",
           remove = TRUE) %>%
  separate(other,
           into = c("Chromosome", "Cluster"),
           sep = ",",
           extra = "merge",
           remove = TRUE) %>%
  mutate(Cluster = gsub(")", "", Cluster)) %>%
    arrange(Chromosome) %>%
    mutate(Chromosome = gsub("Chr", "", Chromosome)) %>%
    mutate(Chromosome = case_when(!is.na(as.numeric(Chromosome)) ~ as.numeric(Chromosome),
                                  Chromosome == "X" ~ 23,
                                  Chromosome == "Y" ~ 24,
                                  Chromosome == "X|Y" ~ 25,
                                  Chromosome == "MT" ~ 26,
                                  .default = NA));

t_p4c_soma_lm_singleaptamer_t21only_vif
t_p4c_soma_lm_singleaptamer_t21only_tidy
t_p4c_soma_lm_singleaptamer_t21only_glance
t_p4c_soma_lm_singleaptamer_t21only_augment
```

#### Save output to files with prefix 't_p4c_soma_lm_singleaptamer_'
```{r}
setwd(dir.ddata)
fwrite(t_p4c_soma_lm_singleaptamer_t21only_vif, "t_p4f_soma_lm_singleaptamer_t21only_vif.csv.gz")
fwrite(t_p4c_soma_lm_singleaptamer_t21only_tidy, "t_p4f_soma_lm_singleaptamer_t21only_tidy.csv.gz")
fwrite(t_p4c_soma_lm_singleaptamer_t21only_augment, "t_p4f_soma_lm_singleaptamer_t21only_augment.csv.gz")
fwrite(t_p4c_soma_lm_singleaptamer_t21only_glance, "t_p4f_soma_lm_singleaptamer_t21only_glance.csv.gz")
```


#### Create listing of aptamers significantly associated with age or sex in T21s 
```{r}
l_p4c_soma_lm_singleaptamer_t21only_sigage_up <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(term == "Age" & PadjBH<0.10 & Zlog10FC > 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);
l_p4c_soma_lm_singleaptamer_t21only_sigage_down <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(term == "Age" & PadjBH<0.10 & Zlog10FC < 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);

l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(term == "Female" & PadjBH<0.10 & Zlog10FC > 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);

l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(term == "Female" & PadjBH<0.10 & Zlog10FC < 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);

modelc_aptamer_lists <- list(l_p4c_soma_lm_singleaptamer_t21only_sigage_up,
                             l_p4c_soma_lm_singleaptamer_t21only_sigage_down,
                             l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up,
                             l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down);
names(modelc_aptamer_lists) <- c("l_p4c_soma_lm_singleaptamer_t21only_sigage_up",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigage_down",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down")

for ( i in 1:length(modelc_aptamer_lists) ){
  setwd(dir.ddata)
  write.table(modelc_aptamer_lists[[i]], paste0(names(modelc_aptamer_lists)[[i]], ".txt"),
              quote = FALSE,
              row.names = FALSE,
              col.names = FALSE)
  }
```

#### Create listings of aptamers significantly associated with each model term and output to files with prefix 'l_p4c_soma_lm_singleaptamer_modelc_'
```{r}
l_p4c_soma_lm_singleaptamer_t21only_sigage_up <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(grepl("T21*Age", model_formula)==TRUE & term == "Age" & PadjBH<0.10 & Zlog10FC > 0) %>%
  dplyr::select(Aptamer) %>%
  unique();
l_p4c_soma_lm_singleaptamer_t21only_sigage_down <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(grepl("T21*Age", model_formula)==TRUE & term == "Age" & PadjBH<0.10 & Zlog10FC < 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);

l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(grepl("T21*Age", model_formula)==TRUE & term == "Female" & PadjBH<0.10 & Zlog10FC > 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);
l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down <- t_p4c_soma_lm_singleaptamer_t21only_tidy %>%
  filter(grepl("T21*Age", model_formula)==TRUE & term == "Female" & PadjBH<0.10 & Zlog10FC < 0) %>%
  dplyr::select(Aptamer) %>%
  unique() %>%
  `rownames<-`(NULL);

modelc_aptamer_lists <- list(l_p4c_soma_lm_singleaptamer_t21only_sigage_up,
                             l_p4c_soma_lm_singleaptamer_t21only_sigage_down,
                             l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up,
                             l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down);
names(modelc_aptamer_lists) <- c("l_p4c_soma_lm_singleaptamer_t21only_sigage_up",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigfge_down",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigfemale_up",
                                 "l_p4c_soma_lm_singleaptamer_t21only_sigfemale_down")

for ( i in 1:length(modelc_aptamer_lists) ){
  setwd(dir.ddata)
  write.table(modelc_aptamer_lists[[i]], paste0(names(modelc_aptamer_lists)[[i]], ".txt"),
              quote = FALSE,
              row.names = FALSE,
              col.names = FALSE)
  }
```
